---
title: "permutation"
author: "Meng"
date: "2024-06-10"
output: html_document
---

```{r setup, include=FALSE}
library(arrow)
library(dplyr)
library(ggplot2)
library(patchwork)
library(qvalue)
data <- "/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/eqtl/"
```


# permutations, empirical p-values for phenotypes
phenotype-level summary statistics with empirical p-values, enabling calculation of genome-wide FDR

## histogram of empirical p values for phenotypes
use beta-approximated p-values pval_beta
```{r}
ML <- read.table(paste0(data, "ML.cis_qtl.txt"), header = T) # 14658
FL <- read.table(paste0(data, "FL.cis_qtl.txt"), header = T) # 14733
MP <- read.table(paste0(data, "MP.cis_qtl.txt"), header = T) # 13794

hist(ML$pval_beta, nclass = 20)
hist(FL$pval_beta, nclass = 20)
hist(MP$pval_beta, nclass = 20)

qval_ML <- qvalue(ML$pval_beta)
qval_FL <- qvalue(FL$pval_beta)
qval_MP <- qvalue(MP$pval_beta)

summary(qval_ML)
p1 <- hist(qval_ML)

summary(qval_FL)
p2 <- hist(qval_FL)

summary(qval_MP)
p3 <- hist(qval_MP)

p1 + p2 + p3 + plo
```
Call:
qvalue(p = ML$pval_beta)

pi0:	0.4482977	

Cumulative number of significant calls:

          <1e-04 <0.001 <0.01 <0.025 <0.05 <0.1    <1
p-value      605   1084  2131   2872  3748 4967 14658
q-value      328    591  1334   1912  2662 4075 14658
local FDR    225    387   803   1129  1526 2203 14649


Call:
qvalue(p = FL$pval_beta)

pi0:	0.4428507	

Cumulative number of significant calls:

          <1e-04 <0.001 <0.01 <0.025 <0.05 <0.1    <1
p-value      720   1208  2234   2972  3835 5083 14733
q-value      417    741  1501   2120  2853 4260 14733
local FDR    264    494   923   1286  1674 2263 14733


Call:
qvalue(p = MP$pval_beta)

pi0:	0.601221	

Cumulative number of significant calls:

          <1e-04 <0.001 <0.01 <0.025 <0.05 <0.1    <1
p-value      468    796  1562   2144  2836 3896 13794
q-value      218    415   785   1103  1511 2192 13794
local FDR    144    268   490    659   877 1213 13794

## extract q values and pval_nominal_threshold
```{r}
calculate_qvalues <- function(res_df, fdr = 0.1, qvalue_lambda = NULL) {
  # Number of phenotypes tested
  num_phenotypes <- nrow(res_df)
  cat(sprintf("Number of phenotypes tested: %d\n", num_phenotypes))
  
  # Determine which p-value column to use
  if ("pval_beta" %in% colnames(res_df) && !all(is.na(res_df$pval_beta))) {
    pval_col <- "pval_beta"
    correlation <- cor(res_df$pval_perm, res_df$pval_beta, use = "complete.obs")
    cat(sprintf("Correlation between Beta-approximated and empirical p-values: %.4f\n", correlation))
  } else {
    pval_col <- "pval_perm"
    cat("WARNING: no beta-approximated p-values found, using permutation p-values instead.\n")
  }
  
  # Calculate q-values
  if (!is.null(qvalue_lambda)) {
    cat(sprintf("Calculating q-values with lambda = %.3f\n", qvalue_lambda))
    qobj <- qvalue(res_df[[pval_col]], lambda = qvalue_lambda)
  } else {
    qobj <- qvalue(res_df[[pval_col]])
  }
  
  res_df$qval <- qobj$qvalues
  pi0 <- qobj$pi0
  cat(sprintf("Proportion of significant phenotypes (1-pi0): %.2f\n", 1 - pi0))
  cat(sprintf("QTL phenotypes @ FDR %.2f: %d\n", fdr, sum(res_df$qval <= fdr)))
  
  # Determine global min(p) significance threshold and calculate nominal p-value threshold for each gene
  if (pval_col == "pval_beta") {
    lb <- sort(res_df[res_df$qval <= fdr, "pval_beta"])
    ub <- sort(res_df[res_df$qval > fdr, "pval_beta"])
    
    if (length(lb) > 0) {
      lb <- tail(lb, 1)
      if (length(ub) > 0) {
        ub <- head(ub, 1)
        pthreshold <- (lb + ub) / 2
      } else {
        pthreshold <- lb
      }
      cat(sprintf("min p-value threshold @ FDR %.2f: %.6g\n", fdr, pthreshold))
      res_df$pval_nominal_threshold <- qbeta(pthreshold, res_df$beta_shape1, res_df$beta_shape2)
    }
  }
  
  return(res_df)
}

ML <- calculate_qvalues(ML)
FL <- calculate_qvalues(FL)
MP <- calculate_qvalues(MP)

```
ML
Number of phenotypes tested: 14658
Correlation between Beta-approximated and empirical p-values: 1.0000
Proportion of significant phenotypes (1-pi0): 0.55
QTL phenotypes @ FDR 0.05: 2662
min p-value threshold @ FDR 0.05: 0.0202564

FL
Number of phenotypes tested: 14733
Correlation between Beta-approximated and empirical p-values: 1.0000
Proportion of significant phenotypes (1-pi0): 0.56
QTL phenotypes @ FDR 0.05: 2853
min p-value threshold @ FDR 0.05: 0.0218665

MP
Number of phenotypes tested: 13794
Correlation between Beta-approximated and empirical p-values: 1.0000
Proportion of significant phenotypes (1-pi0): 0.40
QTL phenotypes @ FDR 0.05: 1511
min p-value threshold @ FDR 0.05: 0.00910948

ML 14658 tested, 4075 (q < 0.1), 2662 (q < 0.05)
FL 14733 tested, 4260 (q < 0.1), 2853 (q < 0.05)
MP 13794 tested, 2192 (q < 0.1), 1511 (q < 0.05)

## extract eGenes
To obtain eGene as FDR 10%, we can collect all genes with qval smaller than 0.1
```{r}
ML_egene <- ML %>% filter(qval < 0.1) # 4075

sum(ML_egene$num_var) # 457888

FL_egene <- FL %>% filter(qval < 0.1) # 4260

MP_egene <- MP %>% filter(qval < 0.1) # 2192

# Venn of egenes
write.csv(MP_egene, file = paste0(data, "MP_egene.csv"), row.names = F, quote = F)
write.csv(ML_egene, file = paste0(data, "ML_egene.csv"), row.names = F, quote = F)
write.csv(FL_egene, file = paste0(data, "FL_egene.csv"), row.names = F, quote = F)

write.table(MP_egene$phenotype_id, file = paste0(data, "MP_egene.txt"), row.names = F, quote = F, col.names = F)
write.table(ML_egene$phenotype_id, file = paste0(data, "ML_egene.txt"), row.names = F, quote = F, col.names = F)
write.table(FL_egene$phenotype_id, file = paste0(data, "FL_egene.txt"), row.names = F, quote = F, col.names = F)
```


## check out the most significant phenotype & its most significant eQTL
```{r}
ML_A1 <- read_parquet(paste0(data,"ML.cis_qtl_pairs.A1.parquet"))
ML_A2 <- read_parquet(paste0(data,"ML.cis_qtl_pairs.A2.parquet"))
ML_A3 <- read_parquet(paste0(data,"ML.cis_qtl_pairs.A3.parquet"))
ML_A4 <- read_parquet(paste0(data,"ML.cis_qtl_pairs.A4.parquet"))

ML_eqtl <- rbind(ML_A1, ML_A2, ML_A3, ML_A4) # 1413340
sum(ML$num_var) # 1413340

TX_paternal_00006106_eqtl <- ML_A1 %>% filter(phenotype_id == "TX_paternal_00006106")

TX_paternal_00006106_eqtl <- TX_paternal_00006106_eqtl %>% arrange(pval_nominal) 

TX_paternal_00006106_eqtl[1,]

TX_paternal_00006106_egene <- ML %>% filter(phenotype_id == "TX_paternal_00006106")

egene <- TX_paternal_00006106_egene %>% select(c("phenotype_id","variant_id","start_distance","af","ma_samples","ma_count","pval_nominal", "slope","slope_se" )) 

egene <- rbind(egene, TX_paternal_00006106_eqtl[1,])

# colnames(TX_paternal_00006106_egene)
# colnames(ML)
```

```{r}
FL_A1 <- read_parquet(paste0(data,"FL.cis_qtl_pairs.A1.parquet"))
FL_A2 <- read_parquet(paste0(data,"FL.cis_qtl_pairs.A2.parquet"))
FL_A3 <- read_parquet(paste0(data,"FL.cis_qtl_pairs.A3.parquet"))
FL_A4 <- read_parquet(paste0(data,"FL.cis_qtl_pairs.A4.parquet"))

FL_eqtl <- rbind(FL_A1, FL_A2, FL_A3, FL_A4) # 1706960
sum(FL$num_var) # 1706960

MP_A1 <- read_parquet(paste0(data,"MP.cis_qtl_pairs.A1.parquet"))
MP_A2 <- read_parquet(paste0(data,"MP.cis_qtl_pairs.A2.parquet"))
MP_A3 <- read_parquet(paste0(data,"MP.cis_qtl_pairs.A3.parquet"))
MP_A4 <- read_parquet(paste0(data,"MP.cis_qtl_pairs.A4.parquet"))

MP_eqtl <- rbind(MP_A1, MP_A2, MP_A3, MP_A4) # 1249621
sum(MP$num_var) # 1249621

```


## extract expression from phenotype file and snp info from VCF for TX_paternal_00006106 A1:274220346
```{r}
data <- "/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/"
normalized_counts_mln <- read.table(paste0(data,"normalized_counts_mln.txt"), header = T)

expr <- normalized_counts_mln %>% filter(gene_id == "TX_paternal_00006106")

geno <- read.table(paste0(data,"A1_274220346.txt"), header = T)

pheno <- expr[,c(5:79)]
geno1 <- geno[,c(5:79)]

eqtl <- rbind(geno1,pheno)

eqtl <- t(eqtl)
eqtl <- as.data.frame(eqtl)

eqtl <- eqtl %>% mutate(id = row.names(eqtl))
colnames(eqtl) <- c("genotype", "phenotype", "id")

eqtl$genotype <- factor(eqtl$genotype)
eqtl$phenotype <- as.numeric(eqtl$phenotype)

ggplot(eqtl, aes(x=genotype, y=phenotype)) + 
  geom_boxplot() + theme_classic()+ geom_jitter(shape=16, position=position_jitter(0.2), colour = "red")
```

```{r}
table(eqtl$genotype)

write.csv(TX_paternal_00006106_eqtl_sig, file = paste0(data,"TX_paternal_00006106_eqtl_sig.csv"), quote = F)

```



## check distance to TSS and number of cis-eQTLs
−log(p)on y-axis and distance to TSS on x-axis. And put a harizontal line indicating the corresponing pval_nominal_threshold of the gene.
```{r}
TX_paternal_00006106_eqtl_full <- inner_join(TX_paternal_00006106_eqtl, ML[,c(1,16:19)], by = "phenotype_id") 

#TX_paternal_00006106_eqtl_sig <- TX_paternal_00006106_eqtl_full %>% filter(pval_nominal < pval_nominal_threshold)

head(TX_paternal_00006106_eqtl_full)


p1 <- ggplot(TX_paternal_00006106_eqtl_full, aes(x=start_distance, y=-log(pval_nominal))) + geom_point() + theme_classic() + geom_hline(yintercept=-log(0.0004534847)) + ggtitle("TX_paternal_00006106")

p2 <- ggplot(TX_paternal_00006106_eqtl_full, aes(x=abs(start_distance), y=-log(pval_nominal))) + geom_point() + theme_classic() + geom_hline(yintercept=-log(0.0004534847)) + ggtitle("TX_paternal_00006106")


gene1 <- ML_eqtl_full %>% filter(phenotype_id == "TX_paternal_00001808")

p3 <- ggplot(gene1, aes(x=abs(start_distance), y=-log(pval_nominal))) + geom_point() + theme_classic() + geom_hline(yintercept=-log(0.0003248906)) + ggtitle("TX_paternal_00001808")

p4 <- ggplot(gene1, aes(x=start_distance, y=-log(pval_nominal))) + geom_point() + theme_classic() + geom_hline(yintercept=-log(0.0003248906)) + ggtitle("TX_paternal_00001808")

p1 + p2 + p3 + p4 + plot_layout(nrow = 2)
```








## extract FDR-corrected significant eQTLs
To obtain cis-eQTL for these eGenes, we can collect all variant/gene pairs with pval_nominal (reported in cis_nominal run) smaller than pval_nominal_threshold.
```{r}
ML_eqtl_full <- inner_join(ML_eqtl, ML[,c(1,16:19)], by = "phenotype_id") # 1413340
#ML_eqtl_full <- inner_join(ML_eqtl, ML_egene[,c(1,16:19)], by = "phenotype_id") # same result
ML_eqtl_sig <- ML_eqtl_full %>% filter(pval_nominal < pval_nominal_threshold) # 23769


FL_eqtl_full <- inner_join(FL_eqtl, FL[,c(1,16:19)], by = "phenotype_id") # 1706960
FL_eqtl_sig <- FL_eqtl_full %>% filter(pval_nominal < pval_nominal_threshold) # 32397


MP_eqtl_full <- inner_join(MP_eqtl, MP[,c(1,16:19)], by = "phenotype_id") # 1249621 
MP_eqtl_sig <- MP_eqtl_full %>% filter(pval_nominal < pval_nominal_threshold) # 12224

```
```{r}
ML_eqtl_sig <- ML_eqtl_sig %>% mutate(maf = ifelse(af <= 0.5, af, 1 - af))
FL_eqtl_sig <- FL_eqtl_sig %>% mutate(maf = ifelse(af <= 0.5, af, 1 - af))
MP_eqtl_sig <- MP_eqtl_sig %>% mutate(maf = ifelse(af <= 0.5, af, 1 - af))


write.table(ML_eqtl_sig, file = paste0(data, "ML_eqtl_sig.txt"), row.names = F, quote = F, sep = "\t")
write.table(FL_eqtl_sig, file = paste0(data, "FL_eqtl_sig.txt"), row.names = F, quote = F, sep = "\t")
write.table(MP_eqtl_sig, file = paste0(data, "MP_eqtl_sig.txt"), row.names = F, quote = F, sep = "\t")

write.csv(ML_eqtl_sig, file = paste0(data, "ML_eqtl_sig.csv"), row.names = F, quote = F)
write.csv(FL_eqtl_sig, file = paste0(data, "FL_eqtl_sig.csv"), row.names = F, quote = F)
write.csv(MP_eqtl_sig, file = paste0(data, "MP_eqtl_sig.csv"), row.names = F, quote = F)

write.csv(ML_eqtl_full, file = paste0(data, "ML_eqtl_full.csv"), row.names = F, quote = F)
write.csv(FL_eqtl_full, file = paste0(data, "FL_eqtl_full.csv"), row.names = F, quote = F)
write.csv(MP_eqtl_full, file = paste0(data, "MP_eqtl_full.csv"), row.names = F, quote = F)
```


## number of SNPs per gene

tested
```{r}
p1 <- ggplot(ML, aes(x=num_var)) + geom_histogram(binwidth=1) + theme_classic() + ggtitle("Male leaf")
p2 <- ggplot(FL, aes(x=num_var)) + geom_histogram(binwidth=1) + theme_classic() + ggtitle("Female leaf")
p3 <- ggplot(MP, aes(x=num_var)) + geom_histogram(binwidth=1) + theme_classic() + ggtitle("Pollen")

p1 + p2 + p3
```


significant
```{r}
p4 <- ggplot(ML_egene, aes(x=num_var)) + geom_histogram(binwidth=1) + theme_classic() + ggtitle("Male leaf")
p5 <- ggplot(FL_egene, aes(x=num_var)) + geom_histogram(binwidth=1) + theme_classic() + ggtitle("Female leaf")
p6 <- ggplot(MP_egene, aes(x=num_var)) + geom_histogram(binwidth=1) + theme_classic() + ggtitle("Pollen")

p1 + p2 + p3 + p4 + p5 + p6 + plot_layout(nrow =2)
```

