---
title: "MAF null"
author: "Meng"
date: "2024-11-13"
output: html_document
---

```{r setup, include=FALSE}
library(arrow)
library(dplyr)
library(ggplot2)
library(patchwork)
data <- "/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/eqtl/"
```

```{r}
get_hist_prop2 <- function(data, binwidth = 0.05, boundary = 0.05, x_limits = c(0.05, 0.5)) {p <- ggplot(data, aes(x = maf)) + geom_histogram(binwidth = binwidth, boundary = boundary) + scale_x_continuous(limits = x_limits)
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y / sum(p_data$y)
p_data <- p_data %>% dplyr::select(x, xmin, xmax, y, y_prop)
return(p_data)}
```


# check the p value distribution of real vs permuted data
```{r}
ML_A1 <- read_parquet(paste0(data,"ML.cis_qtl_pairs.1.parquet"))
ML_A2 <- read_parquet(paste0(data,"ML.cis_qtl_pairs.2.parquet"))
ML_A3 <- read_parquet(paste0(data,"ML.cis_qtl_pairs.3.parquet"))
ML_A4 <- read_parquet(paste0(data,"ML.cis_qtl_pairs.4.parquet"))
ML_eqtl <- rbind(ML_A1, ML_A2, ML_A3, ML_A4) # 1272993

data <- "/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/ml/"
ML_A1 <- read_parquet(paste0(data,"ML_50.cis_qtl_pairs.1.parquet"))
ML_A2 <- read_parquet(paste0(data,"ML_50.cis_qtl_pairs.2.parquet"))
ML_A3 <- read_parquet(paste0(data,"ML_50.cis_qtl_pairs.3.parquet"))
ML_A4 <- read_parquet(paste0(data,"ML_50.cis_qtl_pairs.4.parquet"))
ML_eqtl_50 <- rbind(ML_A1, ML_A2, ML_A3, ML_A4) # 1272993

p1 <- ggplot(ML_eqtl, aes(x=pval_nominal)) + geom_histogram(color="black", fill="white", binwidth = 0.005) + theme_classic() + xlab("p-value")+ggtitle("real")
p2 <- ggplot(ML_eqtl_50, aes(x=pval_nominal)) + geom_histogram(color="black", fill="white", binwidth = 0.005) + theme_classic() + xlab("p-value")+ggtitle("1 permutation")
p1 + p2 + plot_layout(nrow = 2)
```


# ML
## get MAF from real data
```{r}
ML_eqtl_sig <- read.csv(paste0(data, "ML_eqtl_sig_Nov.csv")) # 16674
ML_eqtl_sig <- ML_eqtl_sig %>% group_by(phenotype_id) %>% mutate(count = n()) %>% ungroup() %>% filter(count <= 10) # 5934

set.seed(123)
ML_eqtl_sig_random <- ML_eqtl_sig %>% group_by(phenotype_id) %>% slice_sample(n = 1)  # 2047

ML_eqtl_multigene <- read.csv(paste0(data, "ML_eqtl_multigene.csv"))
ML_eqtl_sig_random_af <- ML_eqtl_sig_random %>% 
    anti_join(ML_eqtl_multigene, by = c("variant_id","maf")) %>%
    dplyr::select(variant_id, maf) # 1953

ML_eqtl_sig_random_af$maf <- round(ML_eqtl_sig_random_af$maf, digits = 3)
p_data <- get_hist_prop2(ML_eqtl_sig_random_af)
write.table(p_data, file = paste0(data, "ML_eqtl_random_real.txt"), row.names = FALSE, quote = FALSE, sep = "\t", col.names = T) 
```

## check number of fake eqtls
```{r}
cnt <- read.table(paste0(data, "ML_fake_cnt.txt"))
summary(cnt$V1)
212/16674
summary(cnt$V2)
182/5934
summary(cnt$V3)
116/1953
```
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  118.0   190.8   212.0   217.9   241.0   418.0 
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  115.0   168.0   182.0   182.9   197.0   261.0 
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
   85.0   108.0   116.0   115.9   123.0   156.0 
   
## plot
```{r}
maf_real <- read.table(paste0(data, "ML_eqtl_random_real.txt"), header = T)
maf_real <- maf_real %>% mutate(x_axis = paste(xmin, xmax, sep = "-"))

maf_null <- read.table(paste0(data, "ML_MAF_null.txt"), header = F)
colnames(maf_null) <- c("x", "xmin","xmax","y", "y_prop")
maf_null <- maf_null %>% mutate(x_axis = paste(xmin, xmax, sep = "-"))

p1 <- ggplot(maf_null, aes(x_axis, y_prop)) + 
    geom_point(fill = NA, colour = "darkgrey", alpha = 0.2)+
    geom_point(data = maf_real, color = "black", fill="black")+ 
    theme_classic()+ labs(x="MAF", y="Proportion", title = "a")+
    theme(axis.text.x = element_text(angle = 45, hjust = 1),text = element_text(size = 10))+
    ylim(0, 0.9)
p1
```

# FL
## get MAF from real data
```{r}
FL_eqtl_sig <- read.csv(paste0(data, "FL_eqtl_sig_Nov.csv")) # 34238
FL_eqtl_sig <- FL_eqtl_sig %>% group_by(phenotype_id) %>% mutate(count = n()) %>% ungroup() %>% filter(count <= 10) # 8090

set.seed(123)
FL_eqtl_sig_random <- FL_eqtl_sig %>% group_by(phenotype_id) %>% slice_sample(n = 1)  # 2677

FL_eqtl_multigene <- read.csv(paste0(data, "FL_eqtl_multigene.csv"))
FL_eqtl_sig_random_af <- FL_eqtl_sig_random %>% 
    anti_join(FL_eqtl_multigene, by = c("variant_id","maf")) %>%
    dplyr::select(variant_id, maf) # 2542

FL_eqtl_sig_random_af$maf <- round(FL_eqtl_sig_random_af$maf, digits = 3)
p_data <- get_hist_prop2(FL_eqtl_sig_random_af)
write.table(p_data, file = paste0(data, "FL_eqtl_random_real.txt"), row.names = FALSE, quote = FALSE, sep = "\t", col.names = T) 
```

## check the number of fake eqtls
```{r}
cnt <- read.table(paste0(data, "FL_fake_cnt.txt"))
summary(cnt$V1)
569/34238
summary(cnt$V2)
450/8090
summary(cnt$V3)
268/2542
```
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  375.0   524.2   569.0   576.0   620.0   933.0 
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  334.0   426.0   450.0   450.5   473.0   559.0
     Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  210.0   257.0   268.0   267.8   278.0   320.0
  
## plot
```{r}
maf_real <- read.table(paste0(data, "FL_eqtl_random_real.txt"), header = T)
maf_real <- maf_real %>% mutate(x_axis = paste(xmin, xmax, sep = "-"))

maf_null <- read.table(paste0(data, "FL_MAF_null.txt"), header = F)
colnames(maf_null) <- c("x", "xmin","xmax","y", "y_prop")
maf_null <- maf_null %>% mutate(x_axis = paste(xmin, xmax, sep = "-"))

p2 <- ggplot(maf_null, aes(x_axis, y_prop)) + 
    geom_point(fill = NA, colour = "darkgrey", alpha = 0.2)+
    geom_point(data = maf_real, color = "black", fill="black")+ 
    theme_classic()+ labs(x="MAF", y="Proportion", title = "b")+
    theme(axis.text.x = element_text(angle = 45, hjust = 1),text = element_text(size = 10))+
    ylim(0, 0.9)
p2
```


# MP
## get MAF from real data
```{r}
MP_eqtl_sig <- read.csv(paste0(data, "MP_eqtl_sig_Nov.csv")) # 9456
MP_eqtl_sig <- MP_eqtl_sig %>% group_by(phenotype_id) %>% mutate(count = n()) %>% ungroup() %>% filter(count <= 10) # 3637

set.seed(123)
MP_eqtl_sig_random <- MP_eqtl_sig %>% group_by(phenotype_id) %>% slice_sample(n = 1)  # 1257

MP_eqtl_multigene <- read.csv(paste0(data, "MP_eqtl_multigene.csv"))
MP_eqtl_sig_random_af <- MP_eqtl_sig_random %>% 
    anti_join(MP_eqtl_multigene, by = c("variant_id","maf")) %>%
    dplyr::select(variant_id, maf) # 1195

MP_eqtl_sig_random_af$maf <- round(MP_eqtl_sig_random_af$maf, digits = 3)
p_data <- get_hist_prop2(MP_eqtl_sig_random_af)
write.table(p_data, file = paste0(data, "MP_eqtl_random_real.txt"), row.names = FALSE, quote = FALSE, sep = "\t", col.names = T) 
```

## check the number of fake eqtls
```{r}
cnt <- read.table(paste0(data, "MP_fake_cnt.txt"))
summary(cnt$V1)
65/9456
summary(cnt$V2)
58/3637
summary(cnt$V3)
38/1195
```
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  28.00   54.00   65.00   69.01   80.00  173.00
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  26.00   50.00   58.00   58.44   66.00  114.00
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  18.00   34.00   38.00   37.91   42.00   63.00   

## plot
```{r}
maf_real <- read.table(paste0(data, "MP_eqtl_random_real.txt"), header = T)
maf_real <- maf_real %>% mutate(x_axis = paste(xmin, xmax, sep = "-"))

maf_null <- read.table(paste0(data, "MP_MAF_null.txt"), header = F)
colnames(maf_null) <- c("x", "xmin","xmax","y", "y_prop")
maf_null <- maf_null %>% mutate(x_axis = paste(xmin, xmax, sep = "-"))

p3 <- ggplot(maf_null, aes(x_axis, y_prop)) + 
    geom_point(fill = NA, colour = "darkgrey", alpha = 0.2)+
    geom_point(data = maf_real, color = "black", fill="black")+ 
    theme_classic()+ labs(x="MAF", y="Proportion", title = "c")+
    theme(axis.text.x = element_text(angle = 45, hjust = 1),text = element_text(size = 10))+
    ylim(0, 0.9)
p3
```

# plot together
```{r}
p1 + p2 + p3
ggsave(paste0(data,"MAF_null_Jun3.pdf"),width = 7.7, height = 3.2, units = c("in"))
```