---
title: "phenotype correlation on autosomes"
author: "Meng"
date: "2025-05-22"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(patchwork)
data <- "/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/eqtl/"
```

## between sex
```{r}
normalized_counts_ln <- read.table("/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/normalized_counts_ln_auto.bed", header = T, check.names = F) # 14992
col <- as.data.frame(colnames(normalized_counts_ln))
male <- normalized_counts_ln[,c(5:73)]
female <- normalized_counts_ln[,c(74:147)]

male <- male[, order(names(male))]
female <- female[, order(names(female))]

write.table(colnames(male), file=paste0(data, "m.txt"), quote = F, row.names = F)
write.table(colnames(female), file=paste0(data, "f.txt"), quote = F, row.names = F)
# keep the same inds
# remove from male: "14eM", "37fM", "40bM", "45dM", "67fM"
# remove from female: "24aF", "27cF", "35bF", "53cF", "5dF", "64aF", "72dF", "73dF", "7cF", "8aF"
cols_to_remove <- c("14eM", "37fM", "40bM", "45dM", "67fM")
male <- male %>% dplyr::select(-all_of(cols_to_remove)) # 64 pairs

cols_to_remove <- c("24aF", "27cF", "35bF", "53cF", "5dF", "64aF", "72dF", "73dF", "7cF", "8aF")
female <- female %>% dplyr::select(-all_of(cols_to_remove)) # 64 pairs

row.names(male) <- normalized_counts_ln[,4]
row.names(female) <- normalized_counts_ln[,4]

cor(as.numeric(male[3, ]), as.numeric(female[3, ]), method = "pearson")

sex_r <- sapply(1:nrow(male), function(i) cor(as.numeric(male[i, ]), as.numeric(female[i, ]), method = "pearson"))
names(sex_r) <- rownames(male)

sex_r_df <- data.frame(
  phenotype_id = rownames(male),
  pearson_r = sex_r
)

# label wether the gene is an egene in either ML or FL
egene_ml <- read.table(paste0(data, "ML_egene_eqtl.txt"))
egene_fl <- read.table(paste0(data, "FL_egene_eqtl.txt"))
colnames(egene_ml) <- "phenotype_id"
colnames(egene_fl) <- "phenotype_id"

sex_r_df$Category <- ifelse(
  sex_r_df$phenotype_id %in% egene_ml$phenotype_id | sex_r_df$phenotype_id %in% egene_fl$phenotype_id,
  "eGene",
  "Not eGene")
table(sex_r_df$Category)
sex_r_df$Category <- factor(sex_r_df$Category)
levels(sex_r_df$Category)
sex_r_df$Category <- factor(sex_r_df$Category, levels = c("Not eGene", "eGene"))
```
eGene Not eGene 
     4493     10499

### plot the distribution of r
```{r}
p1 <- ggplot(sex_r_df, aes(x = pearson_r, colour = Category)) + geom_histogram(binwidth = 0.05, position = "identity", alpha = 0.5, fill = NA) + theme_classic() + labs(title = "a", x = "Pearson's r", y = "Count") + xlim(-1, 1) + ylim(0,1550) + scale_color_brewer(palette="Dark2")
p1
```


save output
```{r}
write.csv(sex_r_df, file=paste0(data, "sex_r_df.csv"), quote = F, row.names = F)
sex_r_df <- read.csv(paste0(data, "sex_r_df.csv"))
```



## between leaf/pollen
has to be normalized together
```{r}
normalized_counts_pln <- read.table("/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/normalized_counts_pln.bed", header = T, check.names = F)

# keep only autosome
normalized_counts_pln <- normalized_counts_pln %>% filter(chrom != "PAR" & chrom != "X" & chrom != "Y") # 16411

pollen <- normalized_counts_pln %>% select(ends_with("MP"))
leaf <- normalized_counts_pln %>% select(ends_with("ML"))
colnames(pollen)
colnames(leaf)

row.names(pollen) <- normalized_counts_pln[,4]
row.names(leaf) <- normalized_counts_pln[,4]


cor(as.numeric(pollen[3, ]), as.numeric(leaf[3, ]), method = "pearson")

pl_r <- sapply(1:nrow(pollen), function(i) cor(as.numeric(pollen[i, ]), as.numeric(leaf[i, ]), method = "pearson"))
names(pl_r) <- rownames(pollen)

pl_r_df <- data.frame(
  phenotype_id = rownames(pollen),
  pearson_r = pl_r)

# label wether the gene is an egene in either ML or FL
egene_mp <- read.table(paste0(data, "MP_egene_eqtl.txt"))
colnames(egene_mp) <- "phenotype_id"

pl_r_df$Category <- ifelse(
  pl_r_df$phenotype_id %in% egene_ml$phenotype_id | pl_r_df$phenotype_id %in% egene_mp$phenotype_id,
  "eGene",
  "Not eGene")
table(pl_r_df$Category)
pl_r_df$Category <- factor(pl_r_df$Category)
levels(pl_r_df$Category)
pl_r_df$Category <- factor(pl_r_df$Category, levels = c("Not eGene", "eGene"))
```
eGene Not eGene 
     3469     12942

### plot the distribution of r
```{r}
p2 <- ggplot(pl_r_df, aes(x = pearson_r, colour = Category)) + geom_histogram(binwidth = 0.05, position = "identity", alpha = 0.5, fill=NA) + theme_classic() + labs(title = "b", x = "Pearson's r", y = "Count") + xlim(-1, 1) + ylim(0,1550) + scale_color_brewer(palette="Dark2")
p2
```


plot together
```{r}
p1 + p2 + plot_layout(guides = 'collect')
ggsave(paste0(data, "phenotype_correlation_pearson.pdf"), height = 3.6, width = 7.5, units = "in")
```

save output
```{r}
write.csv(pl_r_df, file=paste0(data, "pl_r_df.csv"), quote = F, row.names = F)
pl_r_df <- read.csv(paste0(data, "pl_r_df.csv"))
```


## plot the other way
```{r}
egene_sex <- sex_r_df %>% filter(Category == "eGene")
egene_pl <- pl_r_df %>% filter(Category == "eGene")
egene_sex$Category <- "Sex"
egene_pl$Category <- "Life-stage"
egene_r <- rbind(egene_sex, egene_pl)
egene_r$Category <- as.factor(egene_r$Category)
levels(egene_r$Category)
table(egene_r$Category)
egene_median <- egene_r %>% group_by(Category) %>% summarise(median = median(pearson_r))
wilcox.test(data = egene_r, pearson_r ~ Category)
```
Life-stage        Sex 
      3469       4493 

Wilcoxon rank sum test with continuity correction

data:  pearson_r by Category
W = 4380880, p-value < 2.2e-16
alternative hypothesis: true location shift is not equal to 0

```{r}
p3 <- ggplot(egene_r, aes(x = pearson_r, fill = Category, colour = Category)) + geom_histogram(binwidth = 0.05, position = "identity", alpha = 0.1) + theme_classic() + labs(title = "b", x = expression("Pearson's " * italic(r)), y = "Count") + xlim(-1, 1) + scale_fill_brewer(palette="Dark2") + scale_color_brewer(palette="Dark2") + geom_vline(data=egene_median, aes(xintercept=median, color=Category), linetype="dashed")
p3
```

```{r}
pl <- pl_r_df
sex <- sex_r_df
sex$Category <- "Sex"
pl$Category <- "Life-stage"
all_r <- rbind(sex, pl)
all_r$Category <- as.factor(all_r$Category)
levels(all_r$Category)
all_median <- all_r %>% group_by(Category) %>% summarise(median = median(pearson_r))
wilcox.test(data = all_r, pearson_r ~ Category)
```
Wilcoxon rank sum test with continuity correction

data:  pearson_r by Category
W = 61665951, p-value < 2.2e-16
alternative hypothesis: true location shift is not equal to 0

```{r}
p4 <- ggplot(all_r, aes(x = pearson_r, fill = Category, colour = Category)) + geom_histogram(binwidth = 0.05, position = "identity", alpha = 0.1) + theme_classic() + labs(title = "a", x = expression("Pearson's " * italic(r)), y = "Count") + xlim(-1, 1) + scale_fill_brewer(palette="Dark2") + scale_color_brewer(palette="Dark2")+ geom_vline(data=egene_median, aes(xintercept=median, color=Category), linetype="dashed")
p4
```

```{r}
p4 + p3 + plot_layout(guides = 'collect')
ggsave(paste0(data, "Fig2.pdf"), height = 3.4, width = 7.5, units = "in")
```

extrat top sex correlated eGenes
```{r}
test <- egene_sex %>% filter(pearson_r > 0.8)

top100 <- egene_sex[order(egene_sex$pearson_r, decreasing = TRUE), ][1:100, ]

top05 <-  egene_sex %>% filter(pearson_r > 0.5) # 462

close0 <-  egene_sex %>% filter(pearson_r > -0.1 & pearson_r < 0.1) # 348



write.csv(ML_eqtl_sig_min, file = paste0(data, "ML_eqtl_sig_min.csv"), row.names = F, quote = F)
write.csv(FL_eqtl_sig_min, file = paste0(data, "FL_eqtl_sig_min.csv"), row.names = F, quote = F)
ML_eqtl_sig_min <- read.csv(paste0(data, "ML_eqtl_sig_min.csv"))
FL_eqtl_sig_min <- read.csv(paste0(data, "FL_eqtl_sig_min.csv"))

ml <- inner_join(top100, ML_eqtl_sig_min, by = "phenotype_id")
fl <- inner_join(top100, FL_eqtl_sig_min, by = "phenotype_id")

ml_high <- inner_join(top05, ML_eqtl_sig_min, by = "phenotype_id") # 289
ml_low <- inner_join(close0, ML_eqtl_sig_min, by = "phenotype_id") # 135

fl_high <- inner_join(top05, FL_eqtl_sig_min, by = "phenotype_id") # 388
fl_low <- inner_join(close0, FL_eqtl_sig_min, by = "phenotype_id") # 248


median(ml_high$maf)
median(ml_low$maf)
median(fl_high$maf)
median(fl_low$maf)

median(ml$maf)
median(fl$maf)

p<-ggplot(fl_high, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.117

p1 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("a")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10))  + expand_limits(x=0.5) + geom_vline(xintercept=0.1016949, color = "red",linetype="dashed")


p<-ggplot(fl_low, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.117

p2 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("a")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10))  + expand_limits(x=0.5) + geom_vline(xintercept=0.1016949, color = "red",linetype="dashed")
p1 + p2
```

```{r}
get_hist_prop2 <- function(data, binwidth = 0.05, boundary = 0.05, x_limits = c(0.05, 0.5)) {p <- ggplot(data, aes(x = maf)) + geom_histogram(binwidth = binwidth, boundary = boundary) + scale_x_continuous(limits = x_limits)
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y / sum(p_data$y)
p_data <- p_data %>% dplyr::select(x, xmin, xmax, y, y_prop)
return(p_data)}

ml_high$maf <- round(ml_high$maf, digits = 3)
p_data <- get_hist_prop2(ml_high)


```

```{r}
wilcox.test(ml_high$maf, ml_low$maf)
wilcox.test(fl_high$maf, fl_low$maf)
```

