---
title: "male leaf vs female leaf differential expresion"
author: "Meng"
date: "2024-05-22"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library("DESeq2")
library('RColorBrewer')
library("stringr")
library(VennDiagram)
library(rattle)
data <- "/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/"
```

## male leaf vs female leaf
already run deseq2 for A and PAR in phenotype_L.Rmd
can rerun for sex chroms here
```{r}
# unwanted samples already removed
readcnt_leaf <- read.table(paste0(data,"readcnt_leaf_sept.txt"), header = T, check.names = F)
col<-data.frame(colnames(readcnt_leaf))
readcnt_leaf <- readcnt_leaf %>%  mutate(m_mean = rowMeans(readcnt_leaf[,c(2:70)] )) %>% mutate(f_mean = rowMeans(readcnt_leaf[,c(71:144)] ))
# remove genes with low read counts
readcnt_leaf <- readcnt_leaf %>% filter(m_mean >= 5 | f_mean >= 5) # 37659 -> 19494

readcnt_matrix<-as.matrix(readcnt_leaf[ , -c(1,145,146)]) 
rownames(readcnt_matrix) <- readcnt_leaf[ , 1]
str(readcnt_matrix)
meta_data <- data.frame(colnames(readcnt_matrix))

colnames(meta_data) <- "id"
sex <- rep(NA, length(meta_data$id))
sex[grep("M", meta_data$id)] <- "male"
sex[grep("F", meta_data$id)] <- "female"
meta_data <- data.frame(meta_data, sex)
rownames(meta_data) <- meta_data[,1]
head(meta_data)

# run deseq2
dds <- DESeqDataSetFromMatrix(countData = readcnt_matrix, 
                              colData = meta_data,
                              design = ~ sex)
dds <- DESeq(dds)
resultsNames(dds) # "sex_male_vs_female"
res <- results(dds)
head(res)
summary(res)
result <- res[complete.cases(res),]
nrow(res) #  19494
nrow(result) # 19492

DE_sex<-as.data.frame(result)
DE_sex <- DE_sex %>% mutate(gene = rownames(DE_sex))
DE_sex <- DE_sex[,c(7,1:6)]
DE_sex <- inner_join(DE_sex, gff, by = "gene")
write.table(DE_sex, file = paste0(data,"DE_sex_full.txt"), row.names = FALSE, quote = F)
```

## expression PCA including sex chroms
```{r}
vsdata <- vst(dds, blind=FALSE)
#+ geom_text(aes(label=name),vjust=2)
p2 <- plotPCA(vsdata, intgroup=c("sex"))  
p2
#plotPCA(vsdata, intgroup=c("sex"))  + geom_text(aes(label=name),vjust=2)
```



## DE genes
```{r}
DE_sex <- read.table(paste0(data,"DE_sex_full.txt"), header = T)
table(DE_sex$chrom_new)
head(DE_sex) 
DE_sex <- DE_sex %>% filter(padj<0.05) # 5636

results_male <- data.frame(chrom = c("A1", "A2", "A3", "A4", "PAR", "X", "Y"), log2FC_1 = NA, log2FC_2 = NA, log2FC_3 = NA)
results_female <- data.frame(chrom = c("A1", "A2", "A3", "A4", "PAR", "X", "Y"), log2FC_1 = NA, log2FC_2 = NA, log2FC_3 = NA)

log2_fold_changes <- c(0, 1, 2)
for (i in seq_along(log2_fold_changes)) {
  DE_sex_male <- DE_sex %>% filter(log2FoldChange > log2_fold_changes[i] & padj < 0.05)
  DE_sex_male$chrom_new <- factor(DE_sex_male$chrom_new, levels=c("A1", "A2", "A3", "A4", "PAR", "X", "Y"))
  tablem <- as.data.frame(table(DE_sex_male$chrom_new))
  results_male[1:7, paste0("log2FC_", i)] <- tablem[c(1:7), 2]  
  
  DE_sex_female <- DE_sex %>% filter(log2FoldChange < -log2_fold_changes[i] & padj < 0.05)
  DE_sex_female$chrom_new <- factor(DE_sex_female$chrom_new, levels=c("A1", "A2", "A3", "A4", "PAR", "X", "Y"))
  tablef <- as.data.frame(table(DE_sex_female$chrom_new))
  results_female[1:7, paste0("log2FC_", i)] <- tablef[c(1:7), 2]  
}
write.csv(results_male, paste0(data,"DE_male_cnt.csv"), row.names = FALSE)
write.csv(results_female, paste0(data,"DE_female_cnt.csv"), row.names = FALSE)
```


## Venn diagram
### all genes, FC = 2
```{r}
DE_tissue <- read.table(paste0(data,"DE_tissue_full.txt"), header = T)
DE_sex_m <-  DE_sex %>% filter(log2FoldChange > 1 & padj < 0.05) # 1091
DE_sex_f <-  DE_sex %>% filter(log2FoldChange < -1 & padj < 0.05) # 405
DE_tissue_pollen <- DE_tissue %>% filter(log2FoldChange > 1 & padj < 0.05) # 6702
DE_tissue_leaf <-  DE_tissue %>% filter(log2FoldChange < -1 & padj < 0.05) # 9947

v1<-DE_tissue_pollen$gene
v2<-DE_tissue_leaf$gene
v3<-DE_sex_f$gene
v4<-DE_sex_m$gene

venn.diagram(
  x = list(v1, v2, v3, v4),
  category.names = c("pollen-bias" , "leaf-bias" , "female-bias", "male-bias"),
  filename = 'venn_FC2_p05.png',
  output=TRUE,
  # Set names
  cat.cex = 1,
  cat.fontface = "bold")

write.table(v1, file = paste0(data, "pollen.txt"), row.names = F, quote = F, col.names = F)
write.table(v2, file = paste0(data, "leaf.txt"), row.names = F, quote = F, col.names = F)
write.table(v3, file = paste0(data, "f.txt"), row.names = F, quote = F, col.names = F)
write.table(v4, file = paste0(data, "m.txt"), row.names = F, quote = F, col.names = F)

pm <- inner_join(DE_tissue_pollen,DE_sex_m, by = c("gene","chrom", "start", "end", "chrom_new"))
table(pm$chrom_new)

pf <- inner_join(DE_tissue_pollen,DE_sex_f, by = c("gene","chrom", "start", "end", "chrom_new"))
table(pf$chrom_new)
```
 A1  A2  A3  A4 PAR   X   Y 
  9   9   6   1   5   1 269 

 A1  A2  A3 PAR   X   Y 
  4   1   1   1  63   3  
  
### all genes, FC = 4
```{r}
DE_sex_m <-  DE_sex %>% filter(log2FoldChange > 2 & padj < 0.05) # 755
DE_sex_f <-  DE_sex %>% filter(log2FoldChange < -2 & padj < 0.05) # 5
DE_tissue_pollen <- DE_tissue %>% filter(log2FoldChange > 2 & padj < 0.05) # 5205
DE_tissue_leaf <-  DE_tissue %>% filter(log2FoldChange < -2 & padj < 0.05) # 6297

v1<-DE_tissue_pollen$gene
v2<-DE_tissue_leaf$gene
v3<-DE_sex_f$gene
v4<-DE_sex_m$gene

venn.diagram(
  x = list(v1, v2, v3, v4),
  category.names = c("pollen-bias" , "leaf-bias" , "female-bias", "male-bias"),
  filename = 'venn_FC4_p05.png',
  output=TRUE,
  # Set names
  cat.cex = 1,
  cat.fontface = "bold")

write.table(v1, file = paste0(data, "pollen_FC4.txt"), row.names = F, quote = F, col.names = F)
write.table(v2, file = paste0(data, "leaf_FC4.txt"), row.names = F, quote = F, col.names = F)
write.table(v3, file = paste0(data, "f_FC4.txt"), row.names = F, quote = F, col.names = F)
write.table(v4, file = paste0(data, "m_FC4.txt"), row.names = F, quote = F, col.names = F)
```

### all genes, FC = 1
```{r}
DE_sex_m <-  DE_sex %>% filter(log2FoldChange > 0 & padj < 0.05) # 3321
DE_sex_f <-  DE_sex %>% filter(log2FoldChange < 0 & padj < 0.05) # 2315
DE_tissue_pollen <- DE_tissue %>% filter(log2FoldChange > 0 & padj < 0.05) # 7991
DE_tissue_leaf <-  DE_tissue %>% filter(log2FoldChange < 0 & padj < 0.05) # 11831

v1<-DE_tissue_pollen$gene
v2<-DE_tissue_leaf$gene
v3<-DE_sex_f$gene
v4<-DE_sex_m$gene

venn.diagram(
  x = list(v1, v2, v3, v4),
  category.names = c("pollen-bias" , "leaf-bias" , "female-bias", "male-bias"),
  filename = 'venn_FC1_p05.png',
  output=TRUE,
  # Set names
  cat.cex = 1,
  cat.fontface = "bold")

# write.table(v1, file = paste0(data, "pollen_FC4.txt"), row.names = F, quote = F, col.names = F)
# write.table(v2, file = paste0(data, "leaf_FC4.txt"), row.names = F, quote = F, col.names = F)
# write.table(v3, file = paste0(data, "f_FC4.txt"), row.names = F, quote = F, col.names = F)
# write.table(v4, file = paste0(data, "m_FC4.txt"), row.names = F, quote = F, col.names = F)
```



## contigency test
FC 2
male female
all, 1091, 405
pollen, 300, 73

```{r}
data <- matrix(c(1091, 405, 300, 73), nrow = 2, byrow = TRUE)

rownames(data) <- c("all", "pollen")
colnames(data) <- c("male", "female")

table_data <- as.table(data)

fisher.test(table_data)
```
Fisher's Exact Test for Count Data

data:  table_data
p-value = 0.00281
alternative hypothesis: true odds ratio is not equal to 1
95 percent confidence interval:
 0.4882361 0.8724196
sample estimates:
odds ratio 
 0.6556372
 
```{r}
# data <- matrix(c(1581, 826, 466, 389), nrow = 2, byrow = TRUE)
# 
# rownames(data) <- c("all", "leaf")
# colnames(data) <- c("male", "female")
# 
# table_data <- as.table(data)
# 
# fisher.test(table_data)
```

 X     A
Female-bias
Sex-bias


Location Female Biased, all genes tested for DE
X              a  378             b 1970
Autosome       c   10            d 14992

Location Male Biased, all genes tested for DE
Y              a  1026             b 1454
Autosome       c   50            d 14992
```{r}
data <- matrix(c(370, 10, 1970, 14992), nrow = 2,
            dimnames = list(Location = c("X", "Autosome"),
            Status = c("FemaleBiased", "All")))

chisq.test(data)$p.value # 0

data <- matrix(c(1026, 50, 1454, 14992), nrow = 2,
            dimnames = list(Location = c("Y", "Autosome"),
            Status = c("MaleBiased", "All")))

chisq.test(data)$p.value # 0

```

```{r}
mosaicplot(table_data)
```

