---
title: "conditionally independent eQTLs"
author: "Meng"
date: "2024-11-08"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(patchwork)
library(tidyr)
library(car)
data <- "/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/eqtl/"
ML_eqtl_sig <- read.csv(paste0(data, "ML_eqtl_sig_Nov.csv"))
FL_eqtl_sig <- read.csv(paste0(data, "FL_eqtl_sig_Nov.csv"))
MP_eqtl_sig <- read.csv(paste0(data, "MP_eqtl_sig_Nov.csv"))

ML <- read.table(paste0(data, "ML.cis_qtl.txt"), header = T) # 14777
FL <- read.table(paste0(data, "FL.cis_qtl.txt"), header = T) # 14564
MP <- read.table(paste0(data, "MP.cis_qtl.txt"), header = T) # 13785
ML_egene <- ML %>% filter(qval < 0.1) # 2425
FL_egene <- FL %>% filter(qval < 0.1) # 3445
MP_egene <- MP %>% filter(qval < 0.1) # 1481
```


# ML
2792 eqtls for 2425 genes (eGenes)
```{r}
ml <- read.table(paste0(data, "ML.cis_independent_qtl.txt"), header = T) # 2792
length(unique(ml$phenotype_id)) # 2425
ml <- ml %>% mutate(maf = ifelse(af <= 0.5, af, 1 - af)) # for maf ananlysis later

egene <- as.data.frame(unique(ML_eqtl_sig$phenotype_id)) 
colnames(egene) <- "phenotype_id"
ml <- inner_join(ml, egene, by = "phenotype_id") # 2788
length(unique(ml$phenotype_id)) # 2421
ml1 <- ml %>% filter(rank == 1) # 2403

# filter for primary eQTLs
ml1 <- ml %>%
  group_by(phenotype_id) %>% 
  filter(rank == min(rank))
# manually update the top rank for each eGene to rank = 1
ml1$rank <- 1 # 2421
# filter for secondary eQTLs
ml2 <- anti_join(ml, ml1, by = c("phenotype_id", "variant_id")) # 367

# most significant eQTL in each eGene
test1 <- inner_join(ML_egene, ml1, by = c("phenotype_id", "variant_id")) %>% select(phenotype_id, variant_id, rank) # 2339
# rank = 1 but not the most significant eQTL
test2 <- anti_join(ml1, ML_egene, by = c("phenotype_id", "variant_id")) # 82

# write output
write.csv(ml1, file = paste0(data, "ml1.csv"), row.names = F, quote = F)
write.csv(ml2, file = paste0(data, "ml2.csv"), row.names = F, quote = F)
```


## compare primary and secondary eQTLs
```{r}
mean(abs(ml1$slope)) # 1.314509
mean(abs(ml2$slope)) # 1.151421

ml1$category <- "primary"
ml2$category <- "secondary"
ml <- rbind(ml1, ml2)
leveneTest(data = ml, abs(slope) ~ category)
t.test(data = ml, abs(slope) ~ category, var.equal = T)

mean(abs(ml1$start_distance)) # 7922.758
mean(abs(ml2$start_distance)) # 8697.025
leveneTest(data = ml, abs(start_distance) ~ category)
t.test(data = ml, abs(start_distance) ~ category, var.equal = T)

median(ml1$maf) # 0.101695
median(ml2$maf) # 0.0967742
wilcox.test(ml1$maf, ml2$maf)
```

Two Sample t-test

data:  abs(slope) by category
t = 6.7868, df = 2786, p-value = 1.395e-11
alternative hypothesis: true difference in means between group primary and group secondary is not equal to 0
95 percent confidence interval:
 0.1159692 0.2102069
sample estimates:
  mean in group primary mean in group secondary 
               1.314509                1.151421 

Two Sample t-test

data:  abs(start_distance) by category
t = -2.3367, df = 2786, p-value = 0.01952
alternative hypothesis: true difference in means between group primary and group secondary is not equal to 0
95 percent confidence interval:
 -1423.9821  -124.5511
sample estimates:
  mean in group primary mean in group secondary 
               7922.758                8697.025 

Wilcoxon rank sum test with continuity correction

data:  ml1$maf and ml2$maf
W = 466404, p-value = 0.1232
alternative hypothesis: true location shift is not equal to 0

# FL
```{r}
fl <- read.table(paste0(data, "FL.cis_independent_qtl.txt"), header = T) # 4167
length(unique(fl$phenotype_id)) # 3445
fl <- fl %>% mutate(maf = ifelse(af <= 0.5, af, 1 - af)) # for maf ananlysis later

egene <- as.data.frame(unique(FL_eqtl_sig$phenotype_id)) 
colnames(egene) <- "phenotype_id"
fl <- inner_join(fl, egene, by = "phenotype_id") # 4164
length(unique(fl$phenotype_id)) # 3442
fl1 <- fl %>% filter(rank == 1) # 3407

# filter for primary eQTLs
fl1 <- fl %>%
  group_by(phenotype_id) %>% 
  filter(rank == min(rank))
# manually update the top rank for each eGene to rank = 1
fl1$rank <- 1 # 3442
# filter for secondary eQTLs
fl2 <- anti_join(fl, fl1, by = c("phenotype_id", "variant_id")) # 722

# most significant eQTL in each eGene
test1 <- inner_join(FL_egene, fl1, by = c("phenotype_id", "variant_id")) %>% select(phenotype_id, variant_id, rank) # 3327
# rank = 1 but not the most significant eQTL
test2 <- anti_join(fl1, FL_egene, by = c("phenotype_id", "variant_id")) # 115

# write output
write.csv(fl1, file = paste0(data, "fl1.csv"), row.names = F, quote = F)
write.csv(fl2, file = paste0(data, "fl2.csv"), row.names = F, quote = F)
```


## compare primary and secondary eQTLs
```{r}
mean(abs(fl1$slope)) # 1.230636
mean(abs(fl2$slope)) # 1.069681

fl1$category <- "primary"
fl2$category <- "secondary"
fl <- rbind(fl1, fl2)
leveneTest(data = fl, abs(slope) ~ category)
t.test(data = fl, abs(slope) ~ category, var.equal = T)

mean(abs(fl1$start_distance)) # 8081.14
mean(abs(fl2$start_distance)) # 8725.489
leveneTest(data = fl, abs(start_distance) ~ category)
t.test(data = fl, abs(start_distance) ~ category, var.equal = T)

median(fl1$maf) # 0.106061
median(fl2$maf) # 0.1
wilcox.test(fl1$maf, fl2$maf)
```
Two Sample t-test

data:  abs(slope) by category
t = 12.496, df = 4162, p-value < 2.2e-16
alternative hypothesis: true difference in means between group primary and group secondary is not equal to 0
95 percent confidence interval:
 0.1357025 0.1862069
sample estimates:
  mean in group primary mean in group secondary 
               1.230636                1.069681
               
Two Sample t-test

data:  abs(start_distance) by category
t = -2.6222, df = 4162, p-value = 0.008769
alternative hypothesis: true difference in means between group primary and group secondary is not equal to 0
95 percent confidence interval:
 -1126.1139  -162.5833
sample estimates:
  mean in group primary mean in group secondary 
               8081.140                8725.489 

Wilcoxon rank sum test with continuity correction

data:  fl1$maf and fl2$maf
W = 1268679, p-value = 0.3739
alternative hypothesis: true location shift is not equal to 0

# MP
```{r}
mp <- read.table(paste0(data, "MP.cis_independent_qtl.txt"), header = T) # 1640
length(unique(mp$phenotype_id)) # 1481
mp <- mp %>% mutate(maf = ifelse(af <= 0.5, af, 1 - af)) # for maf ananlysis later

egene <- as.data.frame(unique(MP_eqtl_sig$phenotype_id)) 
colnames(egene) <- "phenotype_id"
mp <- inner_join(mp, egene, by = "phenotype_id") # 1634
length(unique(mp$phenotype_id)) # 1475
mp1 <- mp %>% filter(rank == 1) # 1472

# filter for primary eQTLs
mp1 <- mp %>%
  group_by(phenotype_id) %>% 
  filter(rank == min(rank))
# manually update the top rank for each eGene to rank = 1
mp1$rank <- 1 # 1475
# filter for secondary eQTLs
mp2 <- anti_join(mp, mp1, by = c("phenotype_id", "variant_id")) # 159

# most significant eQTL in each eGene
test1 <- inner_join(MP_egene, mp1, by = c("phenotype_id", "variant_id")) %>% select(phenotype_id, variant_id, rank) # 1448
# rank = 1 but not the most significant eQTL
test2 <- anti_join(mp1, MP_egene, by = c("phenotype_id", "variant_id")) # 27

# write output
write.csv(mp1, file = paste0(data, "mp1.csv"), row.names = F, quote = F)
write.csv(mp2, file = paste0(data, "mp2.csv"), row.names = F, quote = F)
```


## compare primary and secondary eQTLs
```{r}
mean(abs(mp1$slope)) # 1.364946
mean(abs(mp2$slope)) # 1.179296

mp1$category <- "primary"
mp2$category <- "secondary"
mp <- rbind(mp1, mp2)
leveneTest(data = mp, abs(slope) ~ category)
t.test(data = mp, abs(slope) ~ category, var.equal = F)

mean(abs(mp1$start_distance)) # 7799.113
mean(abs(mp2$start_distance)) # 8294.013
leveneTest(data = mp, abs(start_distance) ~ category)
t.test(data = mp, abs(start_distance) ~ category, var.equal = T)

median(mp1$maf) # 0.104839
median(mp2$maf) # 0.0964912
wilcox.test(mp1$maf, mp2$maf)
```
	Welch Two Sample t-test

data:  abs(slope) by category
t = 7.7296, df = 188.12, p-value = 6.302e-13
alternative hypothesis: true difference in means between group primary and group secondary is not equal to 0
95 percent confidence interval:
 0.1382708 0.2330301
sample estimates:
  mean in group primary mean in group secondary 
               1.364946                1.179296
               
Two Sample t-test

data:  abs(start_distance) by category
t = -0.98655, df = 1632, p-value = 0.324
alternative hypothesis: true difference in means between group primary and group secondary is not equal to 0
95 percent confidence interval:
 -1478.8359   489.0358
sample estimates:
  mean in group primary mean in group secondary 
               7799.113                8294.013 
               
Wilcoxon rank sum test with continuity correction

data:  mp1$maf and mp2$maf
W = 124145, p-value = 0.2234
alternative hypothesis: true location shift is not equal to 0