---
title: "sex interaction"
author: "Meng"
date: "2024-07-03"
output: html_document
---

```{r setup, include=FALSE}
library(arrow)
library(dplyr)
library(ggplot2)
library(patchwork)
library(tidyr)
data <- "/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/sex/"
```


## permutation
/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/eqtl/sex.cis_qtl.txt
```{r}
sex_genes <- read.table(paste0(data, "sex.cis_qtl.txt"), header = T) # 14755

hist(sex_genes$pval_beta, nclass = 20)
qval_sex <- qvalue(sex_genes$pval_beta)
summary(qval_sex)
hist(qval_sex)
```


```{r}
#sex_genes <- calculate_qvalues(sex_genes)

#sex_egene <- sex_genes %>% filter(qval < 0.05) # 7522
sex_egene <- sex_genes %>% filter(qval < 0.1) # 9743
```
Number of phenotypes tested: 14549
Correlation between Beta-approximated and empirical p-values: 1.0000
Proportion of significant phenotypes (1-pi0): 0.78
QTL phenotypes @ FDR 0.05: 7522
min p-value threshold @ FDR 0.05: 0.117321


## get siginicant eQTLs without interaction

```{r}
sex_A1 <- read_parquet(paste0(data,"sex.cis_qtl_pairs.1.parquet"))
sex_A2 <- read_parquet(paste0(data,"sex.cis_qtl_pairs.2.parquet"))
sex_A3 <- read_parquet(paste0(data,"sex.cis_qtl_pairs.3.parquet"))
sex_A4 <- read_parquet(paste0(data,"sex.cis_qtl_pairs.4.parquet"))
sex_eqtl <- rbind(sex_A1, sex_A2, sex_A3, sex_A4) # 1368353
sum(sex_genes$num_var) # 1368353

sex_eqtl_full <- inner_join(sex_eqtl, sex_genes[,c(1,16:19)], by = "phenotype_id") #
sex_eqtl_sig <- sex_eqtl_full %>% filter(pval_nominal < pval_nominal_threshold) #  116128

nrow(sex_egene) # 9743
length(unique(sex_eqtl_sig$phenotype_id)) # 10344
length(unique(sex_eqtl_sig$variant_id)) # 109449

write.csv(sex_eqtl_sig, file = paste0(data, "sex_eqtl_sig.csv"), row.names = F, quote = F)
```
MAF
```{r}
# MAF
sex_eqtl_sig <- sex_eqtl_sig %>% mutate(maf = ifelse(af <= 0.5, af, 1 - af))
write.csv(sex_eqtl_sig, file = paste0(data, "sex_eqtl_sig.csv"), row.names = F, quote = F)

# unique
sex_eqtl_sig_af <- ML_eqtl_sig %>% select(variant_id, af, maf) %>% distinct()

# most sigificant
ML_eqtl_sig_min <- ML_eqtl_sig %>%
  group_by(phenotype_id) %>% 
  filter(pval_nominal == min(pval_nominal)) # 3131

# random

```


## examine the eqtl with the lowest pval_gi
A1:159008371 TX_paternal_00002596
ref G alt A
```{r}
# data <- "/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/"
normalized_counts_ln <- read.table(paste0(data,"normalized_counts_ln.bed"), header = T)
expr <- normalized_counts_ln %>% filter(gene_id == "TX_paternal_00002596")
geno <- read.table(paste0(data,"A1_159008371.vcf"), header = T)

pheno <- expr[,c(5:153)]
geno1 <- geno[,c(5:153)]
eqtl <- rbind(geno1,pheno)
eqtl <- as.data.frame(t(eqtl))

eqtl <- eqtl %>% mutate(id = row.names(eqtl))
colnames(eqtl) <- c("genotype", "phenotype", "id")

eqtl$genotype <- factor(eqtl$genotype)
levels(eqtl$genotype)
eqtl$phenotype <- as.numeric(eqtl$phenotype)

sex <- rep(NA, length(eqtl$id))
sex[grep("M", eqtl$id)] <- "M"
sex[grep("F", eqtl$id)] <- "F"
eqtl <- data.frame(eqtl, sex)

ggplot(eqtl, aes(x=genotype, y=phenotype, colour=sex)) + 
  geom_boxplot(position=position_dodge2(0.85)) + 
  geom_jitter(shape=16, position=position_jitterdodge(0.2), alpha=0.3) + 
  theme_classic()

write.table(eqtl, file = paste0(data, "A1_159008371.eqtl"), quote = F)
```


```{r}
eqtl <- eqtl %>% filter(genotype == "0/0" | genotype == "0/1")
levels(eqtl$genotype) <-c ("./.","GG", "GA","GG")

p1 <- ggplot(eqtl, aes(x=sex, y=phenotype, colour=genotype)) + 
  geom_boxplot(position=position_dodge2(0.85)) + 
  geom_jitter(shape=16, position=position_jitterdodge(0.2), alpha=0.3) + 
  theme_classic()+ theme(text = element_text(size = 15))

p2 <- ggplot(eqtl, aes(x=genotype, y=phenotype, colour=sex)) + 
  geom_boxplot(position=position_dodge2(0.85)) + 
  geom_jitter(shape=16, position=position_jitterdodge(0.2), alpha=0.3) + 
  theme_classic() + theme(text = element_text(size = 18))+ scale_colour_manual(values = c("red", "blue"))+labs(x="Genotype", y="Expression")
p2 
```



