---
title: "addtional info"
author: "Meng"
date: "2024-06-20"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(patchwork)
data <- "/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/eqtl/"
```

need to produce files from eqtl_permutation.Rmd eg ML_eqtl_full
```{r}
ML_eqtl_sig <-read.csv(paste0(data, "ML_eqtl_sig.csv"))
FL_eqtl_sig <-read.csv(paste0(data, "FL_eqtl_sig.csv"))
MP_eqtl_sig <-read.csv(paste0(data, "MP_eqtl_sig.csv"))

```


# number of significant eqtl per gene
```{r}

ml_table <- ML_eqtl_sig %>% group_by(phenotype_id) %>%
  dplyr::summarise(
    count = length(variant_id),
    mean_tss = mean(abs(start_distance)),
    median_tss = median(abs(start_distance)))

fl_table <- FL_eqtl_sig %>% group_by(phenotype_id) %>%
  dplyr::summarise(
    count = length(variant_id),
    mean_tss = mean(abs(start_distance)),
    median_tss = median(abs(start_distance)))

mp_table <- MP_eqtl_sig %>% group_by(phenotype_id) %>%
  dplyr::summarise(
    count = length(variant_id),
    mean_tss = mean(abs(start_distance)),
    median_tss = median(abs(start_distance)))


ml_table$category <- "Male_leaf"
fl_table$category <- "Female_leaf"
mp_table$category <- "Pollen"

write.table(ml_table, file = paste0(data, "ml_table.txt"), row.names = F, quote = F, col.names = F)
write.table(fl_table, file = paste0(data, "fl_table.txt"), row.names = F, quote = F, col.names = F)
write.table(mp_table, file = paste0(data, "mp_table.txt"), row.names = F, quote = F, col.names = F)


median(ml_table$count)
median(fl_table$count)
median(mp_table$count)

tablec <- rbind(ml_table,fl_table,mp_table)

ggplot(tablec, aes(x=category, y=count)) + 
  geom_boxplot() + theme_classic() + ylim(1,10)
```


# distance to TSS and number of cis-eQTLs
−log(p)on y-axis and distance to TSS on x-axis. And put a harizontal line indicating the corresponing pval_nominal_threshold of the gene.
```{r}
median(ml_table$median_tss) # 7331
median(fl_table$median_tss) # 7587
median(mp_table$median_tss) # 7056.5

mean(ml_table$median_tss) # 8089.743
mean(fl_table$median_tss) # 8251.179
mean(mp_table$median_tss) # 8034.049

```



```{r}
gene1 <- ML_eqtl_full %>% filter(phenotype_id == "TX_paternal_00029650")
gene2 <- FL_eqtl_full %>% filter(phenotype_id == "TX_paternal_00030127")
gene3 <- MP_eqtl_full %>% filter(phenotype_id == "TX_paternal_00006330")


p1 <- ggplot(gene1, aes(x=start_distance, y=-log(pval_nominal))) + geom_point(alpha = 0.6) + theme_classic() + geom_hline(yintercept=-log(0.0011044), color = "red") + labs(title = "a")+ theme(text = element_text(size = 10))
p2 <- ggplot(gene2, aes(x=start_distance, y=-log(pval_nominal))) + geom_point(alpha = 0.6) + theme_classic() + geom_hline(yintercept=-log(0.00101239), color = "red") + labs(title = "b")+ theme(text = element_text(size = 10))
p3 <- ggplot(gene3, aes(x=start_distance, y=-log(pval_nominal))) + geom_point(alpha = 0.6) + theme_classic() + geom_hline(yintercept=-log(0.000283815), color = "red") + labs(title = "c")+ theme(text = element_text(size = 10))

p1 + p2 + p3 
```




# eqtls that affect multiple genes in the same tissue
```{r}
length(unique(FL_eqtl_sig$variant_id)) # 34176
FL_eqtl_multigene <- FL_eqtl_sig %>% 
     group_by(variant_id) %>%
     filter(n()>1) # 1952

length(unique(ML_eqtl_sig$variant_id)) # 16722
ML_eqtl_multigene <- ML_eqtl_sig %>% 
     group_by(variant_id) %>%
     filter(n()>1) # 1058

length(unique(MP_eqtl_sig$variant_id)) # 9616
MP_eqtl_multigene <- MP_eqtl_sig %>% 
     group_by(variant_id) %>%
     filter(n()>1) # 434

p1 <- ggplot(ML_eqtl_multigene, aes(x=maf)) + geom_histogram(binwidth=0.01) + theme_classic() + ggtitle("Male leaf")
p2 <- ggplot(FL_eqtl_multigene, aes(x=maf)) + geom_histogram(binwidth=0.01) + theme_classic() + ggtitle("Female leaf")
p3 <- ggplot(MP_eqtl_multigene, aes(x=maf)) + geom_histogram(binwidth=0.01) + theme_classic() + ggtitle("Male pollen")

p1 + p2 + p3
```


```{r}
wilcox.test(ML_eqtl_multigene$maf, MP_eqtl_multigene$maf)
```



# venn diagram of significant eQTLs (should be unique)
```{r}
v1<-ML_eqtl_sig$variant_id
v2<-FL_eqtl_sig$variant_id
v3<-MP_eqtl_sig$variant_id

write.table(v1, file = paste0(data, "ml_variant.txt"), row.names = F, quote = F, col.names = F)
write.table(v2, file = paste0(data, "fl_variant.txt"), row.names = F, quote = F, col.names = F)
write.table(v3, file = paste0(data, "mp_variant.txt"), row.names = F, quote = F, col.names = F)

# v1 <- unique(v1) # 22643
# 
# venn.diagram(
#   x = list(v1, v2, v3),
#   category.names = c("male leaf" , "female leaf" , "pollen"),
#   filename = 'venn_diagramm_eQTL.png',
#   output=TRUE,
#   # Set names
#   cat.cex = 1,
#   cat.fontface = "bold")
```
