---
title: "effect size plots"
author: "Meng"
date: "2024-10-18"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(patchwork)
data <- "/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/eqtl/"
```

# Effect size plots
```{r}
ML_eqtl_sig_min <- read.csv(paste0(data, "ML_eqtl_sig_min.csv"))
FL_eqtl_sig_min <- read.csv(paste0(data, "FL_eqtl_sig_min.csv"))
MP_eqtl_sig_min <- read.csv(paste0(data, "MP_eqtl_sig_min.csv"))

# important: need to generate FL_eqtl, FL etc from eqtl_permutation.Rmd  
ML_eqtl_full <- inner_join(ML_eqtl, ML[,c(1,16:19)], by = "phenotype_id") # 1272993
FL_eqtl_full <- inner_join(FL_eqtl, FL[,c(1,16:19)], by = "phenotype_id") # 1687298
MP_eqtl_full <- inner_join(MP_eqtl, MP[,c(1,16:19)], by = "phenotype_id") # 1182102

ML_eqtl_full <- ML_eqtl_full %>% mutate(category = ifelse(pval_nominal < pval_nominal_threshold & qval < 0.1, "Significant", "Not_Significant"))
ML_eqtl_sig_min <- ML_eqtl_sig_min %>% select(phenotype_id, variant_id)
ML_eqtl_sig_min$category2 <- "min"
ML_eqtl_full <- left_join(ML_eqtl_full, ML_eqtl_sig_min, by = c("phenotype_id","variant_id"))
ML_eqtl_full <- ML_eqtl_full %>% mutate(category = ifelse(!is.na(category2), category2, category)) %>% dplyr::select(phenotype_id, variant_id, slope, category)


FL_eqtl_full <- FL_eqtl_full %>% mutate(category = ifelse(pval_nominal < pval_nominal_threshold & qval < 0.1, "Significant", "Not_Significant"))
FL_eqtl_sig_min <- FL_eqtl_sig_min %>% select(phenotype_id, variant_id)
FL_eqtl_sig_min$category2 <- "min"
FL_eqtl_full <- left_join(FL_eqtl_full, FL_eqtl_sig_min, by = c("phenotype_id","variant_id"))
FL_eqtl_full <- FL_eqtl_full %>% mutate(category = ifelse(!is.na(category2), category2, category)) %>% dplyr::select(phenotype_id, variant_id, slope, category)


MP_eqtl_full <- MP_eqtl_full %>% mutate(category = ifelse(pval_nominal < pval_nominal_threshold & qval < 0.1, "Significant", "Not_Significant"))
MP_eqtl_sig_min <- MP_eqtl_sig_min %>% select(phenotype_id, variant_id)
MP_eqtl_sig_min$category2 <- "min"
MP_eqtl_full <- left_join(MP_eqtl_full, MP_eqtl_sig_min, by = c("phenotype_id","variant_id"))
MP_eqtl_full <- MP_eqtl_full %>% mutate(category = ifelse(!is.na(category2), category2, category)) %>% dplyr::select(phenotype_id, variant_id, slope, category)

```


## sex
```{r}
ml_fl <- inner_join(ML_eqtl_full, FL_eqtl_full, by = c("phenotype_id","variant_id")) # 679642
ml_fl <- ml_fl %>%  mutate(category = case_when(
    category.x == "min" & category.y == "min" ~ "both",
    category.x == "min" & category.y == "Significant" ~ "male",
    category.x == "min" & category.y == "Not_Significant" ~ "male2",
    category.x == "Significant" & category.y == "min" ~ "female",
    category.x == "Not_Significant" & category.y == "min" ~ "female2"
    ))
ml_fl <- ml_fl[complete.cases(ml_fl),] # 2420
colnames(ml_fl)[3] <- "male"
colnames(ml_fl)[5] <- "female"
write.csv(ml_fl, file = paste0(data, "ml_fl_effectsize.csv"), row.names = F, quote = F)
table(ml_fl$category)

p1 <- ggplot(ml_fl, aes(x = male, y = female, color = category)) + 
    geom_point(data = ml_fl[ml_fl$category != "both", ], size=0.8, alpha = 0.8) +
    geom_point(data = ml_fl[ml_fl$category == "both", ], size=0.8, alpha = 0.8) +
    theme_classic() + labs(x="Male leaf", y="Female leaf", title = "a") + 
    theme(text = element_text(size = 11)) + 
    scale_colour_manual(values = c("black","#B200ED", "#CBC3E3", "#FF6E00", "#FFDBBB")) + 
    geom_vline(xintercept=0, color = "black",linetype="dashed") + 
    geom_hline(yintercept=0, color = "black",linetype="dashed") + 
    xlim(-2.3,2.3) + ylim(-2.3,2.3)
p1
```

both  female female2    male   male2 
    132     221     884     377     806 
    
### 3 categories
```{r}
ml_fl2 <- ml_fl %>%  mutate(category = case_when(
    category.x == "min" & category.y == "min" ~ "both",
    category.x == "min" & category.y == "Significant" ~ "male",
    category.x == "min" & category.y == "Not_Significant" ~ "male",
    category.x == "Significant" & category.y == "min" ~ "female",
    category.x == "Not_Significant" & category.y == "min" ~ "female"
    ))
ml_fl2 <- ml_fl2[complete.cases(ml_fl2),] # 2420
table(ml_fl2$category)

ml_fl2$category <- factor(ml_fl2$category)
levels(ml_fl2$category)

ml_fl2 <- ml_fl2 %>%
  mutate(category = factor(category, levels = c("both", "female", "male")))

    
ggplot(ml_fl2, aes(x = slope.x, y = slope.y, color = category)) + geom_point(data = ml_fl2[ml_fl2$category != "both", ], size=0.8, alpha = 0.8)+ geom_point(data = ml_fl2[ml_fl2$category == "both", ], size=0.8, alpha = 0.8)+ theme_classic() + labs(x="Male leaf", y="Female leaf", title = "a") + theme(text = element_text(size = 11))+ scale_colour_manual(values = c("black", "#CBC3E3", "#FFDBBB"))+ geom_vline(xintercept=0, color = "black",linetype="dashed") + geom_hline(yintercept=0, color = "black",linetype="dashed")

```

 

## lifestage
```{r}
ml_mp <- inner_join(ML_eqtl_full, MP_eqtl_full, by = c("phenotype_id","variant_id")) # 1050294
ml_mp <- ml_mp %>%  mutate(category = case_when(
    category.x == "min" & category.y == "min" ~ "both",
    category.x == "min" & category.y == "Significant" ~ "leaf",
    category.x == "min" & category.y == "Not_Significant" ~ "leaf2",
    category.x == "Significant" & category.y == "min" ~ "pollen",
    category.x == "Not_Significant" & category.y == "min" ~ "pollen2"
    ))
ml_mp <- ml_mp[complete.cases(ml_mp),] # 3077
colnames(ml_mp)[3] <- "leaf"
colnames(ml_mp)[5] <- "pollen"
write.csv(ml_mp, file = paste0(data, "ml_mp_effectsize.csv"), row.names = F, quote = F)

table(ml_mp$category)

p2 <- ggplot(ml_mp, aes(x = leaf, y = pollen, color = category)) + 
    geom_point(data = ml_mp[ml_mp$category != "both", ], size=0.8, alpha = 0.8) +
    geom_point(data = ml_mp[ml_mp$category == "both", ], size=0.8, alpha = 0.8) +
    theme_classic() + labs(x="Male leaf", y="Pollen", title = "b") + 
    theme(text = element_text(size = 11)) + 
    scale_colour_manual(values = c("black","#FF2400", "#FFCCCB", "#0077ec", "#ADD8E6")) + 
    geom_vline(xintercept=0, color = "black",linetype="dashed") + 
    geom_hline(yintercept=0, color = "black",linetype="dashed") + 
    xlim(-2.3,2.3) + ylim(-2.3,2.3)
p2
```
 both    leaf   leaf2  pollen pollen2 
    102      88    1736      94    1057

## plots
```{r}
p1 + p2
ggsave(paste0(data, "Fig3_Jun3.pdf"), height = 3.4, width = 8.3, units = "in")
```



## correlation test
```{r}
cor.test(ml_fl$slope.x, ml_fl$slope.y, method="spearman") 
cor.test(ml_mp$slope.x, ml_mp$slope.y, method="spearman") 
```
Spearman's rank correlation rho

data:  ml_fl\$slope.x and ml_fl$slope.y
S = 662731574, p-value < 2.2e-16
alternative hypothesis: true rho is not equal to 0
sample estimates:
      rho 
0.7194289 


	Spearman's rank correlation rho

data:  ml_mp\$slope.x and ml_mp$slope.y
S = 3173405028, p-value < 2.2e-16
alternative hypothesis: true rho is not equal to 0
sample estimates:
      rho 
0.3464267 



# extract eqtls with discordance 

## between sexes
```{r}
# ml_fl <- read.csv(paste0(data, "ml_fl_effectsize.csv"))
# ml_mp <- read.csv(paste0(data, "ml_mp_effectsize.csv"))

dis_sex <-  ml_fl %>% filter(male * female < 0) # 170
table(dis_sex$category)
dis_sex1 <-  ml_fl %>% filter(male > 0 & female < 0) # 87
dis_sex2 <-  ml_fl %>% filter(male < 0 & female > 0) # 83 

# save egenes to output
write.csv(dis_sex, file = paste0(data, "dis_sex.csv"), row.names = F, quote = F)
write.csv(dis_sex1, file = paste0(data, "dis_sex1.csv"), row.names = F, quote = F)
write.csv(dis_sex2, file = paste0(data, "dis_sex2.csv"), row.names = F, quote = F)
```
female2   male2 
     84      86 

## between leaf and pollen
```{r}
dis_pl <- ml_mp %>% filter(leaf * pollen < 0) # 974
table(dis_pl$category)
dis_pl1 <- ml_mp %>% filter(leaf < 0 & pollen > 0) # 518
dis_pl2 <- ml_mp %>% filter(leaf > 0 & pollen < 0) # 456
table(dis_pl1$category)
table(dis_pl2$category)

# save egenes to output
write.csv(dis_pl, file = paste0(data, "dis_pl.csv"), row.names = F, quote = F)
write.csv(dis_pl1, file = paste0(data, "dis_pl1.csv"), row.names = F, quote = F)
write.csv(dis_pl2, file = paste0(data, "dis_pl2.csv"), row.names = F, quote = F)

# manually check annotation for the 9 eqtls
dis_pl_top <- dis_pl %>% filter(category == "both")
write.csv(dis_pl_top, file = paste0(data, "dis_pl_top.csv"), row.names = F, quote = F)
```
both    leaf   leaf2  pollen pollen2 
      9       5     575       5     380 
 both    leaf   leaf2  pollen pollen2 
      5       2     263       2     246 

   both    leaf   leaf2  pollen pollen2 
      4       3     312       3     134 
      

