---
title: "MAF plots"
author: "Meng"
date: "2024-07-16"
output: html_document
---
```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(patchwork)
data <- "/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/eqtl/"
get_hist_prop <- function(data) {
  p <- ggplot(data, aes(x = maf)) + geom_histogram(binwidth = 0.01,stat="count")
  p_data <- ggplot_build(p)$data[[1]]
  p_data$y_prop <- p_data$y / sum(p_data$y)
  return(p_data)}
```


# MAF for significant cis-eQTL in each tissue

## keep the most significant SNP per gene
```{r}
ML_eqtl_multigene <- read.csv(paste0(data, "ML_eqtl_multigene.csv"))
FL_eqtl_multigene <- read.csv(paste0(data, "FL_eqtl_multigene.csv"))
MP_eqtl_multigene <- read.csv(paste0(data, "MP_eqtl_multigene.csv"))

ML_eqtl_sig_min <- read.csv(paste0(data, "ML_eqtl_sig_min.csv")) # 2421
FL_eqtl_sig_min <- read.csv(paste0(data, "FL_eqtl_sig_min.csv")) # 3442
MP_eqtl_sig_min <- read.csv(paste0(data, "MP_eqtl_sig_min.csv")) # 1475

ML_eqtl_sig_min_af <- ML_eqtl_sig_min %>% 
    anti_join(ML_eqtl_multigene, by = c("variant_id","maf")) %>% 
    dplyr::select(variant_id, maf)  # 2292
FL_eqtl_sig_min_af <- FL_eqtl_sig_min %>% 
    anti_join(FL_eqtl_multigene, by = c("variant_id","maf")) %>%
    dplyr::select(variant_id, maf) # 3234
MP_eqtl_sig_min_af <- MP_eqtl_sig_min %>% 
    anti_join(MP_eqtl_multigene, by = c("variant_id","maf")) %>%
    dplyr::select(variant_id, maf) # 1389
```

## keep a random SNP per gene
```{r}
ML_eqtl_sig_random <- read.csv(paste0(data, "ML_eqtl_sig_random.csv"))
FL_eqtl_sig_random <- read.csv(paste0(data, "FL_eqtl_sig_random.csv"))
MP_eqtl_sig_random <- read.csv(paste0(data, "MP_eqtl_sig_random.csv"))

ML_eqtl_sig_random_af <- ML_eqtl_sig_random %>% 
    anti_join(ML_eqtl_multigene, by = c("variant_id","maf")) %>% 
    dplyr::select(variant_id, maf)  # 2307
FL_eqtl_sig_random_af <- FL_eqtl_sig_random %>% 
    anti_join(FL_eqtl_multigene, by = c("variant_id","maf")) %>%
    dplyr::select(variant_id, maf) # 3265
MP_eqtl_sig_random_af <- MP_eqtl_sig_random %>% 
    anti_join(MP_eqtl_multigene, by = c("variant_id","maf")) %>%
    dplyr::select(variant_id, maf) # 1403
```




# Mann-Whitney U test of MAF of each tissue and prep plots
## most significant
```{r}
median(ML_eqtl_sig_min_af$maf)
median(FL_eqtl_sig_min_af$maf)
median(MP_eqtl_sig_min_af$maf)

IQR(ML_eqtl_sig_min_af$maf)
IQR(FL_eqtl_sig_min_af$maf)
IQR(MP_eqtl_sig_min_af$maf)

wilcox.test(ML_eqtl_sig_min_af$maf, FL_eqtl_sig_min_af$maf)
wilcox.test(ML_eqtl_sig_min_af$maf, MP_eqtl_sig_min_af$maf)
wilcox.test(FL_eqtl_sig_min_af$maf, MP_eqtl_sig_min_af$maf)
```
[1] 0.1016949
[1] 0.1048387
[1] 0.1044776
[1] 0.1018462
[1] 0.1144927
[1] 0.1078977

Wilcoxon rank sum test with continuity correction

data:  ML_eqtl_sig_min_af\$maf and FL_eqtl_sig_min_af$maf
W = 3586978, p-value = 0.04137
alternative hypothesis: true location shift is not equal to 0


	Wilcoxon rank sum test with continuity correction

data:  ML_eqtl_sig_min_af\$maf and MP_eqtl_sig_min_af$maf
W = 1599734, p-value = 0.7995
alternative hypothesis: true location shift is not equal to 0


	Wilcoxon rank sum test with continuity correction

data:  FL_eqtl_sig_min_af\$maf and MP_eqtl_sig_min_af$maf
W = 2327882, p-value = 0.04909
alternative hypothesis: true location shift is not equal to 0


## plots
```{r}
p<-ggplot(ML_eqtl_sig_min_af, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.117

p1 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("a")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.13) + expand_limits(x=0.5) + geom_vline(xintercept=0.1016949, color = "red",linetype="dashed")

p<-ggplot(FL_eqtl_sig_min_af, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.128

p2 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("b")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.13) + expand_limits(x=0.5) + geom_vline(xintercept=0.1048387, color = "red",linetype="dashed")

p<-ggplot(MP_eqtl_sig_min_af, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.128

p3 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("c")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.13) + expand_limits(x=0.5) + geom_vline(xintercept=0.1044776, color = "red",linetype="dashed")
```



## random significant
```{r}
median(ML_eqtl_sig_random_af$maf)
median(FL_eqtl_sig_random_af$maf)
median(MP_eqtl_sig_random_af$maf)

IQR(ML_eqtl_sig_random_af$maf)
IQR(FL_eqtl_sig_random_af$maf)
IQR(MP_eqtl_sig_random_af$maf)

wilcox.test(ML_eqtl_sig_random_af$maf, FL_eqtl_sig_random_af$maf)
wilcox.test(ML_eqtl_sig_random_af$maf, MP_eqtl_sig_random_af$maf)
wilcox.test(FL_eqtl_sig_random_af$maf, MP_eqtl_sig_random_af$maf)
```
[1] 0.09649123
[1] 0.09836066
[1] 0.09821428
[1] 0.09381717
[1] 0.1010929
[1] 0.1008485

Wilcoxon rank sum test with continuity correction

data:  ML_eqtl_sig_random_af\$maf and FL_eqtl_sig_random_af$maf
W = 3649961, p-value = 0.04942
alternative hypothesis: true location shift is not equal to 0


	Wilcoxon rank sum test with continuity correction

data:  ML_eqtl_sig_random_af\$maf and MP_eqtl_sig_random_af$maf
W = 1600030, p-value = 0.5623
alternative hypothesis: true location shift is not equal to 0


	Wilcoxon rank sum test with continuity correction

data:  FL_eqtl_sig_random_af\$maf and MP_eqtl_sig_random_af$maf
W = 2332653, p-value = 0.3169
alternative hypothesis: true location shift is not equal to 0


## plots
```{r}
p<-ggplot(ML_eqtl_sig_random_af, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.127

p4 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("d")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10))+ ylim(0, 0.15) + expand_limits(x=0.5) + geom_vline(xintercept=0.09649123, color = "red",linetype="dashed")

p<-ggplot(FL_eqtl_sig_random_af, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.146

p5 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("e")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10))+ ylim(0, 0.15) + expand_limits(x=0.5) + geom_vline(xintercept=0.09836066, color = "red",linetype="dashed")

p<-ggplot(MP_eqtl_sig_random_af, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.126

p6 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("f")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.15) + expand_limits(x=0.5) + geom_vline(xintercept=0.09821428, color = "red",linetype="dashed")
```


## combine plots
p 4-6 goes to supplementary
```{r}
p1 + p2 + p3 + p4 + p5 + p6 + plot_layout(nrow = 2)
ggsave(paste0(data, "Fig3_mar27.pdf"))
```


# MAF for eqtls in one or two tissues
```{r}
ML_eqtl_sig <-read.csv(paste0(data, "ML_eqtl_sig_Nov.csv"))
MP_eqtl_sig <-read.csv(paste0(data, "MP_eqtl_sig_Nov.csv"))

merged <- bind_rows(ML_eqtl_sig[,1:2] %>% mutate(source = "ML"),
                       MP_eqtl_sig[,1:2] %>% mutate(source = "MP")) %>%
  group_by(phenotype_id, variant_id) %>%
  summarise(count = n(), .groups = "drop")

tissue_n1 <- merged %>% filter(count == 1)
tissue_n2 <- merged %>% filter(count == 2)

all <- rbind(ML_eqtl_sig, MP_eqtl_sig)
all <- all %>% select(phenotype_id, variant_id, maf) %>% distinct()

tissue_n1 <- tissue_n1 %>% left_join(all, by = c("phenotype_id", "variant_id")) # 22738
tissue_n2 <- tissue_n2 %>% left_join(all, by = c("phenotype_id", "variant_id")) # 1696

# plots
p<-ggplot(tissue_n1, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.116
p1 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("a")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.13)+ geom_vline(xintercept=0.1016949, color = "red",linetype="dashed")

p<-ggplot(tissue_n2, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.125
p2 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("b")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.13)+ geom_vline(xintercept=0.1016949, color = "red",linetype="dashed")

p1 + p2
ggsave(paste0(data,"MAF_n1_n2.pdf"))
```

```{r}
median(tissue_n1$maf)
median(tissue_n2$maf)

wilcox.test(tissue_n1$maf, tissue_n2$maf)
```
[1] 0.09836065
[1] 0.1101695

	Wilcoxon rank sum test with continuity correction

data:  tissue_n1\$maf and tissue_n2$maf
W = 18316388, p-value = 0.0005705
alternative hypothesis: true location shift is not equal to 0
