---
title: "addtional info"
author: "Meng"
date: "2024-06-20"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(patchwork)
data <- "/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/eqtl/"
```


# number of SNPs per gene tested
```{r}
p1 <- ggplot(ML, aes(x=num_var)) + geom_histogram(binwidth=1) + theme_classic() + ggtitle("Male leaf")
p2 <- ggplot(FL, aes(x=num_var)) + geom_histogram(binwidth=1) + theme_classic() + ggtitle("Female leaf")
p3 <- ggplot(MP, aes(x=num_var)) + geom_histogram(binwidth=1) + theme_classic() + ggtitle("Pollen")

p1 + p2 + p3
```

# distance to TSS and number of cis-eQTLs
−log(p)on y-axis and distance to TSS on x-axis. And put a harizontal line indicating the corresponing pval_nominal_threshold of the gene.




# eqtls that affect multiple genes in the same tissue
ML: 11882 sig, 11538 unique SNPs, 647 eqtls affect multiple genes
FL: 23944 sig, 23284 unique SNPs, 1244 eqtls affect multiple genes
MP: 6464 sig, 6333 unique SNPs, 261 eqtls affect multiple genes

```{r}
length(unique(FL_eqtl_sig$variant_id)) # 23284
FL_eqtl_multigene <- FL_eqtl_sig %>% 
     group_by(variant_id) %>%
     filter(n()>1) # 1244

length(unique(ML_eqtl_sig$variant_id)) # 11538
ML_eqtl_multigene <- ML_eqtl_sig %>% 
     group_by(variant_id) %>%
     filter(n()>1) # 647

length(unique(MP_eqtl_sig$variant_id)) # 6333
MP_eqtl_multigene <- MP_eqtl_sig %>% 
     group_by(variant_id) %>%
     filter(n()>1) # 261

p1 <- ggplot(ML_eqtl_multigene, aes(x=maf)) + geom_histogram(binwidth=0.01) + theme_classic() + ggtitle("Male leaf")
p2 <- ggplot(FL_eqtl_multigene, aes(x=maf)) + geom_histogram(binwidth=0.01) + theme_classic() + ggtitle("Female leaf")
p3 <- ggplot(MP_eqtl_multigene, aes(x=maf)) + geom_histogram(binwidth=0.01) + theme_classic() + ggtitle("Male pollen")

p1 + p2 + p3
```


```{r}
wilcox.test(ML_eqtl_multigene$maf, MP_eqtl_multigene$maf)
```


# venn diagram of significant eQTLs (should be unique)
```{r}
v1<-ML_eqtl_sig$variant_id
v2<-FL_eqtl_sig$variant_id
v3<-MP_eqtl_sig$variant_id

write.table(v1, file = paste0(data, "ml_variant.txt"), row.names = F, quote = F, col.names = F)
write.table(v2, file = paste0(data, "fl_variant.txt"), row.names = F, quote = F, col.names = F)
write.table(v3, file = paste0(data, "mp_variant.txt"), row.names = F, quote = F, col.names = F)

# v1 <- unique(v1) # 22643
# 
# venn.diagram(
#   x = list(v1, v2, v3),
#   category.names = c("male leaf" , "female leaf" , "pollen"),
#   filename = 'venn_diagramm_eQTL.png',
#   output=TRUE,
#   # Set names
#   cat.cex = 1,
#   cat.fontface = "bold")
```
