---
title: "MAF plots"
author: "Meng"
date: "2024-07-16"
output: html_document
---
```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(patchwork)
data <- "/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/eqtl/"
get_hist_prop <- function(data) {
  p <- ggplot(data, aes(x = maf)) + geom_histogram(binwidth = 0.01,stat="count")
  p_data <- ggplot_build(p)$data[[1]]
  p_data$y_prop <- p_data$y / sum(p_data$y)
  return(p_data)
}
```

# MAF for significant cis-eQTL in each tissue

## keep unique SNPs for MAF
```{r}
ML_eqtl_sig <- read.csv(paste0(data, "ML_eqtl_sig.csv"))
FL_eqtl_sig <- read.csv(paste0(data, "FL_eqtl_sig.csv"))
MP_eqtl_sig <- read.csv(paste0(data, "MP_eqtl_sig.csv"))

# ML_eqtl_sig_af <- ML_eqtl_sig %>% dplyr::select(variant_id, af, maf) %>% distinct()
# FL_eqtl_sig_af <- FL_eqtl_sig %>% dplyr::select(variant_id, af, maf) %>% distinct()
# MP_eqtl_sig_af <- MP_eqtl_sig %>% dplyr::select(variant_id, af, maf) %>% distinct()
```

## keep the most significant SNP per gene
```{r}
ML_eqtl_sig_min <- ML_eqtl_sig %>%
  group_by(phenotype_id) %>% 
  filter(pval_nominal == min(pval_nominal)) # 1899

FL_eqtl_sig_min <- FL_eqtl_sig %>%
  group_by(phenotype_id) %>% 
  filter(pval_nominal == min(pval_nominal)) # 1164

MP_eqtl_sig_min <- MP_eqtl_sig %>%
  group_by(phenotype_id) %>% 
  filter(pval_nominal == min(pval_nominal)) # 2702

ML_eqtl_sig_min_af <- ML_eqtl_sig_min %>% dplyr::select(variant_id, af, maf) %>% distinct()
FL_eqtl_sig_min_af <- FL_eqtl_sig_min %>% dplyr::select(variant_id, af, maf) %>% distinct()
MP_eqtl_sig_min_af <- MP_eqtl_sig_min %>% dplyr::select(variant_id, af, maf) %>% distinct()
```

plots
```{r}
p1 <- ggplot(ML_eqtl_sig_min_af, aes(x=maf)) + geom_histogram(binwidth=0.01) + theme_classic() + ggtitle("Male leaf")+ xlab("MAF")+ theme(text = element_text(size = 10))
p2 <- ggplot(FL_eqtl_sig_min_af, aes(x=maf)) + geom_histogram(binwidth=0.01) + theme_classic() + ggtitle("Female leaf")+ xlab("MAF")+ theme(text = element_text(size = 10))
p3 <- ggplot(MP_eqtl_sig_min_af, aes(x=maf)) + geom_histogram(binwidth=0.01) + theme_classic() + ggtitle("Male pollen")+ xlab("MAF")+ theme(text = element_text(size = 10))
p1 + p2 + p3
```


## keep a random SNP per gene
```{r}
set.seed(123)
MP_eqtl_sig_random <- MP_eqtl_sig %>%
  group_by(phenotype_id) %>% 
  slice_sample(n = 1)

ML_eqtl_sig_random <- ML_eqtl_sig %>%
  group_by(phenotype_id) %>% 
  slice_sample(n = 1)

FL_eqtl_sig_random <- FL_eqtl_sig %>%
  group_by(phenotype_id) %>% 
  slice_sample(n = 1)

ML_eqtl_sig_random_af <- ML_eqtl_sig_random %>% dplyr::select(variant_id, af, maf) %>% distinct()
FL_eqtl_sig_random_af <- FL_eqtl_sig_random %>% dplyr::select(variant_id, af, maf) %>% distinct()
MP_eqtl_sig_random_af <- MP_eqtl_sig_random %>% dplyr::select(variant_id, af, maf) %>% distinct()
```
plots
```{r}
# p4 <- ggplot(ML_eqtl_sig_random_af, aes(x=maf)) + geom_histogram(aes(y=..density..), colour="black", fill="white",binwidth=0.01) + theme_classic() + ggtitle("Male leaf")+ xlab("MAF")+ theme(text = element_text(size = 10))+geom_density(alpha=.2) 
# p5 <- ggplot(FL_eqtl_sig_random_af, aes(x=maf)) + geom_histogram(aes(y=..density..), colour="black", fill="white",binwidth=0.01) + theme_classic() + ggtitle("Female leaf")+ xlab("MAF")+ theme(text = element_text(size = 10))+geom_density(alpha=.2) 
# p6 <- ggplot(MP_eqtl_sig_random_af, aes(x=maf)) + geom_histogram(aes(y=..density..), colour="black", fill="white",binwidth=0.01) + theme_classic() + ggtitle("Male pollen")+ xlab("MAF")+ theme(text = element_text(size = 10))+geom_density(alpha=.2) 
# 
# p4 + p5 + p6
```


# Mann-Whitney U test of MAF of each tissue and prep plots
## most significant
```{r}
median(ML_eqtl_sig_min_af$maf)
median(FL_eqtl_sig_min_af$maf)
median(MP_eqtl_sig_min_af$maf)

IQR(ML_eqtl_sig_min_af$maf)
IQR(FL_eqtl_sig_min_af$maf)
IQR(MP_eqtl_sig_min_af$maf)

wilcox.test(ML_eqtl_sig_min_af$maf, FL_eqtl_sig_min_af$maf)
wilcox.test(ML_eqtl_sig_min_af$maf, MP_eqtl_sig_min_af$maf)
wilcox.test(FL_eqtl_sig_min_af$maf, MP_eqtl_sig_min_af$maf)
```
[1] 0.1171875
[1] 0.1166667
[1] 0.1160714
[1] 0.1158847
[1] 0.1262295
[1] 0.112782
Wilcoxon rank sum test with continuity correction

data:  ML_eqtl_sig_min_af$maf and FL_eqtl_sig_min_af$maf
W = 2503232, p-value = 0.1601
alternative hypothesis: true location shift is not equal to 0


	Wilcoxon rank sum test with continuity correction

data:  ML_eqtl_sig_min_af$maf and MP_eqtl_sig_min_af$maf
W = 1118508, p-value = 0.5759
alternative hypothesis: true location shift is not equal to 0


	Wilcoxon rank sum test with continuity correction

data:  FL_eqtl_sig_min_af$maf and MP_eqtl_sig_min_af$maf
W = 1631618, p-value = 0.06361
alternative hypothesis: true location shift is not equal to 0

```{r}
p<-ggplot(ML_eqtl_sig_min_af, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.096

p1 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("a")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.11)+ geom_vline(xintercept=0.1171875, color = "red",linetype="dashed")

p<-ggplot(FL_eqtl_sig_min_af, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.1065

p2 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("b")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.11)+ geom_vline(xintercept=0.1166667, color = "red",linetype="dashed")

p<-ggplot(MP_eqtl_sig_min_af, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.1065

p3 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("c")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.11)+ geom_vline(xintercept=0.1160714, color = "red",linetype="dashed")
```



## random significant
```{r}
median(ML_eqtl_sig_random_af$maf)
median(FL_eqtl_sig_random_af$maf)
median(MP_eqtl_sig_random_af$maf)

IQR(ML_eqtl_sig_random_af$maf)
IQR(FL_eqtl_sig_random_af$maf)
IQR(MP_eqtl_sig_random_af$maf)

wilcox.test(ML_eqtl_sig_random_af$maf, FL_eqtl_sig_random_af$maf)
wilcox.test(ML_eqtl_sig_random_af$maf, MP_eqtl_sig_random_af$maf)
wilcox.test(FL_eqtl_sig_random_af$maf, MP_eqtl_sig_random_af$maf)
```
[1] 0.112069
[1] 0.1065574
[1] 0.112069
[1] 0.1077907
[1] 0.1182121
[1] 0.115245
	Wilcoxon rank sum test with continuity correction

data:  ML_eqtl_sig_random_af$maf and FL_eqtl_sig_random_af$maf
W = 2523448, p-value = 0.6205
alternative hypothesis: true location shift is not equal to 0


	Wilcoxon rank sum test with continuity correction

data:  ML_eqtl_sig_random_af$maf and MP_eqtl_sig_random_af$maf
W = 1098730, p-value = 0.8707
alternative hypothesis: true location shift is not equal to 0


	Wilcoxon rank sum test with continuity correction

data:  FL_eqtl_sig_random_af$maf and MP_eqtl_sig_random_af$maf
W = 1582935, p-value = 0.4429
alternative hypothesis: true location shift is not equal to 0

```{r}
p<-ggplot(ML_eqtl_sig_random_af, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.111

p4 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("d")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10))+ ylim(0, 0.13)+ geom_vline(xintercept=0.112069, color = "red",linetype="dashed")

p<-ggplot(FL_eqtl_sig_random_af, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.122

p5 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("e")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10))+ ylim(0, 0.13)+ geom_vline(xintercept=0.1065574, color = "red",linetype="dashed")

p<-ggplot(MP_eqtl_sig_random_af, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.111

p6 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("f")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.13)+ geom_vline(xintercept=0.112069, color = "red",linetype="dashed")
```


## combine plots
```{r}
p1 + p2 + p3 + p4 + p5 + p6 + plot_layout(nrow = 2)
```

# write outputs
```{r}
write.csv(ML_eqtl_sig_random_af, file = paste0(data, "ML_eqtl_sig_random_af.csv"), row.names = F, quote = F)
write.csv(FL_eqtl_sig_random_af, file = paste0(data, "FL_eqtl_sig_random_af.csv"), row.names = F, quote = F)
write.csv(MP_eqtl_sig_random_af, file = paste0(data, "MP_eqtl_sig_random_af.csv"), row.names = F, quote = F)

write.csv(ML_eqtl_sig_min_af, file = paste0(data, "ML_eqtl_sig_min_af.csv"), row.names = F, quote = F)
write.csv(FL_eqtl_sig_min_af, file = paste0(data, "FL_eqtl_sig_min_af.csv"), row.names = F, quote = F)
write.csv(MP_eqtl_sig_min_af, file = paste0(data, "MP_eqtl_sig_min_af.csv"), row.names = F, quote = F)
```

