---
title: "FL X and PAR"
author: "Meng"
date: "2024-10-08"
output: html_document
---

```{r setup, include=FALSE}
library(arrow)
library(dplyr)
library(ggplot2)
library(patchwork)
data <- "/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/eqtl/"
```


# PAR, X
```{r}
FL_PAR <- read.table(paste0(data, "FL_PAR.cis_qtl.txt"), header = T) # 1090
FL_X <- read.table(paste0(data, "FL_X.cis_qtl.txt"), header = T) # 1870

hist(FL_PAR$pval_beta, nclass = 20)
hist(FL_X$pval_beta, nclass = 20)

# qval_ML <- qvalue(FL_PAR$pval_beta)
# qval_FL <- qvalue(FL_X$pval_beta)
# summary(qval_ML)
# p1 <- hist(qval_ML)

```


```{r}
FL_PAR <- FL_PAR %>% filter(qval < 0.1) # 111
FL_X <- FL_X %>% filter(qval < 0.1) # 1294

FL_PAR_eqtl <- read_parquet(paste0(data,"FL_PAR.cis_qtl_pairs.6.parquet"))
FL_PAR_eqtl_full <- inner_join(FL_PAR_eqtl, FL_PAR[,c(1,16:19)], by = "phenotype_id") # 11892
FL_PAR_eqtl_sig <- FL_PAR_eqtl_full %>% filter(pval_nominal < pval_nominal_threshold) # 714
FL_PAR_eqtl_sig <-FL_PAR_eqtl_sig %>% mutate(maf = ifelse(af <= 0.5, af, 1 - af))

FL_X_eqtl <- read_parquet(paste0(data,"FL_X.cis_qtl_pairs.5.parquet"))
FL_X_eqtl_full <- inner_join(FL_X_eqtl, FL_X[,c(1,16:19)], by = "phenotype_id") # 184718
FL_X_eqtl_sig <- FL_X_eqtl_full %>% filter(pval_nominal < pval_nominal_threshold) # 16112
FL_X_eqtl_sig <-FL_X_eqtl_sig %>% mutate(maf = ifelse(af <= 0.5, af, 1 - af))

# multigene vs onegene
FL_PAR_eqtl_multigene <- FL_PAR_eqtl_sig %>% 
     group_by(variant_id) %>%
     filter(n()>1) %>%  
    dplyr::select(variant_id, maf) %>% unique() # 39
FL_PAR_eqtl_onegene <- FL_PAR_eqtl_sig %>% 
     group_by(variant_id) %>%
     filter(n()==1) %>%
    dplyr::select(variant_id, maf) # 636

FL_X_eqtl_multigene <- FL_X_eqtl_sig %>% 
     group_by(variant_id) %>%
     filter(n()>1) %>%  
    dplyr::select(variant_id, maf) %>% unique() # 540
FL_X_eqtl_onegene <- FL_X_eqtl_sig %>% 
     group_by(variant_id) %>%
     filter(n()==1) %>%
    dplyr::select(variant_id, maf) # 15027
```


```{r}
set.seed(13)
FL_X_sig_min <- FL_X_eqtl_sig %>%
  group_by(phenotype_id) %>% 
  filter(pval_nominal == min(pval_nominal)) %>% 
  slice_sample(n = 1) # 1294

FL_X_sig_random <- FL_X_eqtl_sig %>%
  group_by(phenotype_id) %>% 
  slice_sample(n = 1)  # 1294

FL_PAR_sig_min <- FL_PAR_eqtl_sig %>%
  group_by(phenotype_id) %>% 
  filter(pval_nominal == min(pval_nominal)) %>% 
  slice_sample(n = 1) # 111

FL_PAR_sig_random <- FL_PAR_eqtl_sig %>%
  group_by(phenotype_id) %>% 
  slice_sample(n = 1)  # 111

FL_X_sig_random_af <- FL_X_sig_random %>% 
    anti_join(FL_X_eqtl_multigene, by = c("variant_id","maf")) %>%
    dplyr::select(variant_id, maf) # 1193
FL_X_sig_min_af <- FL_X_sig_min %>% 
    anti_join(FL_X_eqtl_multigene, by = c("variant_id","maf")) %>%
    dplyr::select(variant_id, maf) # 1187

FL_PAR_sig_random_af <- FL_PAR_sig_random %>% 
    anti_join(FL_PAR_eqtl_multigene, by = c("variant_id","maf")) %>%
    dplyr::select(variant_id, maf) # 103
FL_PAR_sig_min_af <- FL_PAR_sig_min %>% 
    anti_join(FL_PAR_eqtl_multigene, by = c("variant_id","maf")) %>%
    dplyr::select(variant_id, maf) # 103
```


```{r}
# autosomes
FL_eqtl_sig_random <- read.csv(paste0(data, "FL_eqtl_sig_random.csv"))
FL_eqtl_sig_min <- read.csv(paste0(data, "FL_eqtl_sig_min.csv"))
FL_eqtl_multigene <- read.csv(paste0(data, "FL_eqtl_multigene.csv"))
FL_eqtl_sig_random_af <- FL_eqtl_sig_random %>% 
    anti_join(FL_eqtl_multigene, by = c("variant_id","maf")) %>%
    dplyr::select(variant_id, maf) # 3265
FL_eqtl_sig_min_af <- FL_eqtl_sig_min %>% 
    anti_join(FL_eqtl_multigene, by = c("variant_id","maf")) %>%
    dplyr::select(variant_id, maf) # 3234

median(FL_eqtl_sig_min_af$maf)
median(FL_PAR_sig_min_af$maf)
median(FL_X_sig_min_af$maf)
median(FL_eqtl_sig_random_af$maf)
median(FL_PAR_sig_random_af$maf)
median(FL_X_sig_random_af$maf)
```

[1] 0.1048387
[1] 0.111111
[1] 0.09677419
[1] 0.09836066
[1] 0.109375
[1] 0.09027778


```{r}
wilcox.test(FL_eqtl_sig_min_af$maf, FL_PAR_sig_min_af$maf)
wilcox.test(FL_eqtl_sig_min_af$maf, FL_X_sig_min_af$maf)
wilcox.test(FL_PAR_sig_min_af$maf, FL_X_sig_min_af$maf)
```
Wilcoxon rank sum test with continuity correction

data:  FL_eqtl_sig_min_af$maf and FL_PAR_sig_min_af$maf
W = 155288, p-value = 0.242
alternative hypothesis: true location shift is not equal to 0


	Wilcoxon rank sum test with continuity correction

data:  FL_eqtl_sig_min_af$maf and FL_X_sig_min_af$maf
W = 2018591, p-value = 0.008343
alternative hypothesis: true location shift is not equal to 0


	Wilcoxon rank sum test with continuity correction

data:  FL_PAR_sig_min_af$maf and FL_X_sig_min_af$maf
W = 68177, p-value = 0.05204
alternative hypothesis: true location shift is not equal to 0


```{r}
wilcox.test(FL_eqtl_sig_random_af$maf, FL_PAR_sig_random_af$maf)
wilcox.test(FL_eqtl_sig_random_af$maf, FL_X_sig_random_af$maf)
wilcox.test(FL_PAR_sig_random_af$maf, FL_X_sig_random_af$maf)
```

	Wilcoxon rank sum test with continuity correction

data:  FL_eqtl_sig_random_af$maf and FL_PAR_sig_random_af$maf
W = 157698, p-value = 0.2822
alternative hypothesis: true location shift is not equal to 0


	Wilcoxon rank sum test with continuity correction

data:  FL_eqtl_sig_random_af$maf and FL_X_sig_random_af$maf
W = 2067780, p-value = 0.001579
alternative hypothesis: true location shift is not equal to 0


	Wilcoxon rank sum test with continuity correction

data:  FL_PAR_sig_random_af$maf and FL_X_sig_random_af$maf
W = 68824, p-value = 0.04274
alternative hypothesis: true location shift is not equal to 0

# plots

## min
```{r}
# A
p<-ggplot(FL_eqtl_sig_min_af, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.128
p1 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("a")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.185) + expand_limits(x=0.5) + geom_vline(xintercept=0.1048387, color = "red",linetype="dashed")

# PAR
p<-ggplot(FL_PAR_sig_min_af, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.184
p2 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("b")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.185) + expand_limits(x=0.5) + geom_vline(xintercept=0.111111, color = "red",linetype="dashed")

# X
p<-ggplot(FL_X_sig_min_af, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.144
p3 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("c")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.185) + expand_limits(x=0.5) + geom_vline(xintercept=0.09677419, color = "red",linetype="dashed")

p1 + p2 + p3
```

## random
```{r}
# A
p<-ggplot(FL_eqtl_sig_random_af, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.146
p4 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("d")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.214) + expand_limits(x=0.5) + geom_vline(xintercept=0.09836066, color = "red",linetype="dashed")

# PAR
p<-ggplot(FL_PAR_sig_random_af, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.213
p5 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("e")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.214) + expand_limits(x=0.5) + geom_vline(xintercept=0.109375, color = "red",linetype="dashed")

# X
p<-ggplot(FL_X_sig_random_af, aes(x=maf)) + geom_histogram(binwidth=0.01) 
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y/sum(p_data$y) # 0.147
p6 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("f")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.214) + expand_limits(x=0.5) + geom_vline(xintercept=0.09027778, color = "red",linetype="dashed")

p4 + p5 + p6
```
```{r}
p1 + p2 + p3 + p4 + p5 + p6 + plot_layout(nrow = 2)
ggsave(paste0(data, "MAF_PAR_X.pdf"))
```

