---
title: "FL X and PAR"
author: "Meng"
date: "2024-10-08"
output: html_document
---

```{r setup, include=FALSE}
library(arrow)
library(dplyr)
library(ggplot2)
library(patchwork)
data <- "/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/eqtl/"
```


# eQTL mapping on PAR, X in FL
```{r}
FL_PAR <- read.table(paste0(data, "FL_PAR.cis_qtl.txt"), header = T) # 1090
FL_X <- read.table(paste0(data, "FL_X.cis_qtl.txt"), header = T) # 1870

hist(FL_PAR$pval_beta, nclass = 20)
hist(FL_X$pval_beta, nclass = 20)

# qval_ML <- qvalue(FL_PAR$pval_beta)
# qval_FL <- qvalue(FL_X$pval_beta)
# summary(qval_ML)
# p1 <- hist(qval_ML)
```

## identify eQTLs
```{r}
FL_PAR <- FL_PAR %>% filter(qval < 0.1) # 111
FL_X <- FL_X %>% filter(qval < 0.1) # 1294

FL_PAR_eqtl <- read_parquet(paste0(data,"FL_PAR.cis_qtl_pairs.6.parquet"))
FL_PAR_eqtl_full <- inner_join(FL_PAR_eqtl, FL_PAR[,c(1,16:19)], by = "phenotype_id") # 11892
FL_PAR_eqtl_sig <- FL_PAR_eqtl_full %>% filter(pval_nominal < pval_nominal_threshold) # 714
FL_PAR_eqtl_sig <-FL_PAR_eqtl_sig %>% mutate(maf = ifelse(af <= 0.5, af, 1 - af))

FL_X_eqtl <- read_parquet(paste0(data,"FL_X.cis_qtl_pairs.5.parquet"))
FL_X_eqtl_full <- inner_join(FL_X_eqtl, FL_X[,c(1,16:19)], by = "phenotype_id") # 184718
FL_X_eqtl_sig <- FL_X_eqtl_full %>% filter(pval_nominal < pval_nominal_threshold) # 16112
FL_X_eqtl_sig <-FL_X_eqtl_sig %>% mutate(maf = ifelse(af <= 0.5, af, 1 - af))

write.csv(FL_X_eqtl_sig , file = paste0(data, "FL_X_eqtl_sig.csv"), row.names = F, quote = F)
write.csv(FL_PAR_eqtl_sig, file = paste0(data, "FL_PAR_eqtl_sig.csv"), row.names = F, quote = F)
FL_X_eqtl_sig <- read.csv(paste0(data, "FL_X_eqtl_sig.csv"))
FL_PAR_eqtl_sig <- read.csv(paste0(data, "FL_PAR_eqtl_sig.csv"))

# multigene vs onegene
FL_PAR_eqtl_multigene <- FL_PAR_eqtl_sig %>% 
     group_by(variant_id) %>%
     filter(n()>1) %>%  
    dplyr::select(variant_id, maf) %>% unique() # 39
FL_PAR_eqtl_onegene <- FL_PAR_eqtl_sig %>% 
     group_by(variant_id) %>%
     filter(n()==1) %>%
    dplyr::select(variant_id, maf) # 636

FL_X_eqtl_multigene <- FL_X_eqtl_sig %>% 
     group_by(variant_id) %>%
     filter(n()>1) %>%  
    dplyr::select(variant_id, maf) %>% unique() # 540
FL_X_eqtl_onegene <- FL_X_eqtl_sig %>% 
     group_by(variant_id) %>%
     filter(n()==1) %>%
    dplyr::select(variant_id, maf) # 15027
```

## select eQTLs
```{r}
# remove genes with more than 10 eqtls
FL_X_eqtl_sig <- FL_X_eqtl_sig %>% group_by(phenotype_id) %>% mutate(count = n()) %>% ungroup() %>% filter(count <= 10) # 3229
FL_PAR_eqtl_sig <- FL_PAR_eqtl_sig %>% group_by(phenotype_id) %>% mutate(count = n()) %>% ungroup() %>% filter(count <= 10) # 211

set.seed(13)
FL_X_sig_min <- FL_X_eqtl_sig %>%
  group_by(phenotype_id) %>% 
  filter(pval_nominal == min(pval_nominal)) %>% 
  slice_sample(n = 1) # 890
set.seed(13)
FL_X_sig_random <- FL_X_eqtl_sig %>%
  group_by(phenotype_id) %>% 
  slice_sample(n = 1)  # 890
set.seed(13)
FL_PAR_sig_min <- FL_PAR_eqtl_sig %>%
  group_by(phenotype_id) %>% 
  filter(pval_nominal == min(pval_nominal)) %>% 
  slice_sample(n = 1) # 91
set.seed(13)
FL_PAR_sig_random <- FL_PAR_eqtl_sig %>%
  group_by(phenotype_id) %>% 
  slice_sample(n = 1)  # 91

FL_X_sig_random_af <- FL_X_sig_random %>% 
    anti_join(FL_X_eqtl_multigene, by = c("variant_id","maf")) %>%
    dplyr::select(variant_id, maf) # 817
FL_X_sig_min_af <- FL_X_sig_min %>% 
    anti_join(FL_X_eqtl_multigene, by = c("variant_id","maf")) %>%
    dplyr::select(variant_id, maf) # 815

FL_PAR_sig_random_af <- FL_PAR_sig_random %>% 
    anti_join(FL_PAR_eqtl_multigene, by = c("variant_id","maf")) %>%
    dplyr::select(variant_id, maf) # 86
FL_PAR_sig_min_af <- FL_PAR_sig_min %>% 
    anti_join(FL_PAR_eqtl_multigene, by = c("variant_id","maf")) %>%
    dplyr::select(variant_id, maf) # 87

write.csv(FL_PAR_sig_random_af, file = paste0(data, "FL_PAR_sig_random_af.csv"), row.names = F, quote = F)
write.csv(FL_X_sig_random_af, file = paste0(data, "FL_X_sig_random_af.csv"), row.names = F, quote = F)
```

## MAFs
```{r}
# autosomes
FL_eqtl_sig <- read.csv(paste0(data, "FL_eqtl_sig_Nov.csv"))
FL_eqtl_sig <- FL_eqtl_sig %>% group_by(phenotype_id) %>% mutate(count = n()) %>% ungroup() %>% filter(count <= 10) # 8090
set.seed(123)
FL_eqtl_sig_random <- FL_eqtl_sig %>%
  group_by(phenotype_id) %>% 
  slice_sample(n = 1)  # 2677
set.seed(13)
FL_eqtl_sig_min <- FL_eqtl_sig %>%
  group_by(phenotype_id) %>% 
  filter(pval_nominal == min(pval_nominal)) %>% 
  slice_sample(n = 1) # 2677

# FL_eqtl_sig_random <- read.csv(paste0(data, "FL_eqtl_sig_random.csv"))
# FL_eqtl_sig_min <- read.csv(paste0(data, "FL_eqtl_sig_min.csv"))
FL_eqtl_multigene <- read.csv(paste0(data, "FL_eqtl_multigene.csv"))

FL_eqtl_sig_random_af <- FL_eqtl_sig_random %>% 
    anti_join(FL_eqtl_multigene, by = c("variant_id","maf")) %>%
    dplyr::select(variant_id, maf) # 2542
FL_eqtl_sig_min_af <- FL_eqtl_sig_min %>% 
    anti_join(FL_eqtl_multigene, by = c("variant_id","maf")) %>%
    dplyr::select(variant_id, maf) # 2529

median(FL_eqtl_sig_min_af$maf)
median(FL_PAR_sig_min_af$maf)
median(FL_X_sig_min_af$maf)
median(FL_eqtl_sig_random_af$maf)
median(FL_PAR_sig_random_af$maf)
median(FL_X_sig_random_af$maf)
```
[1] 0.09701493
[1] 0.109375
[1] 0.0873016
[1] 0.09375
[1] 0.1069263
[1] 0.08333333

```{r}
wilcox.test(FL_eqtl_sig_min_af$maf, FL_PAR_sig_min_af$maf)
wilcox.test(FL_eqtl_sig_min_af$maf, FL_X_sig_min_af$maf)
wilcox.test(FL_PAR_sig_min_af$maf, FL_X_sig_min_af$maf)
```
	Wilcoxon rank sum test with continuity correction

data:  FL_eqtl_sig_min_af$maf and FL_PAR_sig_min_af$maf
W = 99731, p-value = 0.1378
alternative hypothesis: true location shift is not equal to 0


	Wilcoxon rank sum test with continuity correction

data:  FL_eqtl_sig_min_af$maf and FL_X_sig_min_af$maf
W = 1100346, p-value = 0.003601
alternative hypothesis: true location shift is not equal to 0


	Wilcoxon rank sum test with continuity correction

data:  FL_PAR_sig_min_af$maf and FL_X_sig_min_af$maf
W = 40856, p-value = 0.01933
alternative hypothesis: true location shift is not equal to 0


```{r}
wilcox.test(FL_eqtl_sig_random_af$maf, FL_PAR_sig_random_af$maf)
wilcox.test(FL_eqtl_sig_random_af$maf, FL_X_sig_random_af$maf)
wilcox.test(FL_PAR_sig_random_af$maf, FL_X_sig_random_af$maf)
```
Wilcoxon rank sum test with continuity correction

data:  FL_eqtl_sig_random_af$maf and FL_PAR_sig_random_af$maf
W = 102102, p-value = 0.2979
alternative hypothesis: true location shift is not equal to 0


	Wilcoxon rank sum test with continuity correction

data:  FL_eqtl_sig_random_af$maf and FL_X_sig_random_af$maf
W = 1107988, p-value = 0.003908
alternative hypothesis: true location shift is not equal to 0


	Wilcoxon rank sum test with continuity correction

data:  FL_PAR_sig_random_af$maf and FL_X_sig_random_af$maf
W = 39480, p-value = 0.05877
alternative hypothesis: true location shift is not equal to 0

# MAF plots
```{r}
get_hist_prop <- function(data, binwidth = 0.01) {
  p <- ggplot(data, aes(x = maf)) + geom_histogram(binwidth = binwidth)
  p_data <- ggplot_build(p)$data[[1]]
  p_data$y_prop <- p_data$y / sum(p_data$y)
  return(p_data)}
```

## min
```{r}
# A
p_data <- get_hist_prop(FL_eqtl_sig_min_af) 
p1 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("a")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.21) + expand_limits(x=0.5) + geom_vline(xintercept=0.09701493, color = "red",linetype="dashed")

# PAR
p_data <- get_hist_prop(FL_PAR_sig_min_af) 
p2 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("b")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.21) + expand_limits(x=0.5) + geom_vline(xintercept=0.109375, color = "red",linetype="dashed")

# X
p_data <- get_hist_prop(FL_X_sig_min_af)
p3 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("c")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.21) + expand_limits(x=0.5) + geom_vline(xintercept=0.0873016, color = "red",linetype="dashed")

#p1 + p2 + p3
```

## random
```{r}
# A
p_data <- get_hist_prop(FL_eqtl_sig_random_af)
p4 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("d")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.225) + expand_limits(x=0.5) + geom_vline(xintercept=0.09375, color = "red",linetype="dashed")

# PAR
p_data <- get_hist_prop(FL_PAR_sig_random_af) 
p5 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("e")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.225) + expand_limits(x=0.5) + geom_vline(xintercept=0.1069263, color = "red",linetype="dashed")

# X
p_data <- get_hist_prop(FL_X_sig_random_af)
p6 <- ggplot(p_data, aes(x=x, y=y_prop)) +
  geom_bar(stat="identity", width = 0.01, color="white")  + theme_classic() + ggtitle("f")+ labs(x="MAF", y="Proportion")+ theme(text = element_text(size = 10)) + ylim(0, 0.225) + expand_limits(x=0.5) + geom_vline(xintercept=0.08333333, color = "red",linetype="dashed")

#p4 + p5 + p6
```

```{r}
p1 + p2 + p3 + p4 + p5 + p6 + plot_layout(nrow = 2)
ggsave(paste0(data, "MAF_PAR_X.pdf"))
```


# compare PAR and X in one plot, similar to the null model plot
```{r}
get_hist_prop2 <- function(data, binwidth = 0.05, boundary = 0.05, x_limits = c(0.05, 0.5)) {p <- ggplot(data, aes(x = maf)) + geom_histogram(binwidth = binwidth, boundary = boundary) + scale_x_continuous(limits = x_limits)
p_data <- ggplot_build(p)$data[[1]]
p_data$y_prop <- p_data$y / sum(p_data$y)
p_data <- p_data %>% dplyr::select(x, xmin, xmax, y, y_prop)
return(p_data)}

# A
p_data1  <- read.table(paste0(data, "FL_eqtl_random_real.txt"), header = T)
p_data1  <- p_data1 %>% mutate(x_axis = paste(xmin, xmax, sep = "-"))

FL_PAR_sig_random_af$maf <- round(FL_PAR_sig_random_af$maf, digits = 3)
FL_X_sig_random_af$maf <- round(FL_X_sig_random_af$maf, digits = 3)

p_data2 <- get_hist_prop2(FL_PAR_sig_random_af)
p_data3 <- get_hist_prop2(FL_X_sig_random_af)
p_data2 <- p_data2 %>% mutate(x_axis = paste(xmin, xmax, sep = "-"))
p_data3 <- p_data3 %>% mutate(x_axis = paste(xmin, xmax, sep = "-"))

ggplot(p_data1, aes(x_axis, y_prop)) + 
    geom_point(fill = NA, colour = "black", alpha = 0.9)+
    geom_point(data = p_data2, color = "red", fill=NA, alpha = 0.6)+
    geom_point(data = p_data3, color = "blue", fill=NA, alpha = 0.6)+
    theme_classic()+ labs(x="MAF", y="Proportion", title = "b")+
    theme(axis.text.x = element_text(angle = 45, hjust = 1),text = element_text(size = 10))
```

```{r}
# A
maf_null1 <- read.table(paste0(data, "FL_MAF_null.txt"), header = F)
colnames(maf_null1) <- c("x", "xmin","xmax","y", "y_prop")
maf_null1 <- maf_null1 %>% mutate(x_axis = paste(xmin, xmax, sep = "-"))

# PAR
maf_null2 <- read.table(paste0(data, "FL_PAR_MAF_null.txt"), header = F)
colnames(maf_null2) <- c("x", "xmin","xmax","y", "y_prop")
maf_null2 <- maf_null2 %>% mutate(x_axis = paste(xmin, xmax, sep = "-"))

# X
maf_null3 <- read.table(paste0(data, "FL_X_MAF_null.txt"), header = F)
colnames(maf_null3) <- c("x", "xmin","xmax","y", "y_prop")
maf_null3 <- maf_null3 %>% mutate(x_axis = paste(xmin, xmax, sep = "-"))


ggplot(maf_null1, aes(x_axis, y_prop)) + 
    geom_point(fill = NA, colour = "lightgrey", alpha = 0.9)+
    geom_point(data = maf_null3, color = "lightblue", fill=NA, alpha = 0.6)+
    geom_point(data = p_data1, colour = "black", alpha = 0.9)+
    geom_point(data = p_data2, color = "red", fill=NA, alpha = 0.6)+
    geom_point(data = p_data3, color = "blue", fill=NA, alpha = 0.6)+
    theme_classic()+ labs(x="MAF", y="Proportion", title = "b")+
    theme(axis.text.x = element_text(angle = 45, hjust = 1),text = element_text(size = 10))
ggsave(paste0(data,"FL_A_X_MAF_nul.pdf"))

ggplot(maf_null1, aes(x_axis, y_prop)) + 
    geom_point(fill = NA, colour = "grey", alpha = 0.9)+
    geom_point(data = maf_null3, color = "lightblue", fill=NA, alpha = 0.6)+
    geom_point(data = p_data1, colour = "black", alpha = 0.9)+
    geom_point(data = p_data3, color = "blue", fill=NA, alpha = 0.6)+
    theme_classic()+ labs(x="MAF", y="Proportion", title = "b")+
    theme(axis.text.x = element_text(angle = 45, hjust = 1),text = element_text(size = 10))
ggsave(paste0(data,"FL_A_X_MAF_nul_clean.pdf"))

p1 <- ggplot(maf_null1, aes(x_axis, y_prop)) + 
    geom_point(fill = NA, colour = "darkgrey", alpha = 0.2)+
    geom_point(data = p_data1, color = "black", fill="black")+ 
    theme_classic()+ labs(x="MAF", y="Proportion", title = "a")+
    theme(axis.text.x = element_text(angle = 45, hjust = 1),text = element_text(size = 10))+
    ylim(0, 0.75)
p2 <- ggplot(maf_null3, aes(x_axis, y_prop)) + 
    geom_point(fill = NA, colour = "darkgrey", alpha = 0.2)+
    geom_point(data = p_data3, color = "black", fill="black")+ 
    theme_classic()+ labs(x="MAF", y="Proportion", title = "b")+
    theme(axis.text.x = element_text(angle = 45, hjust = 1),text = element_text(size = 10))+
    ylim(0, 0.75)
p1 + p2
ggsave(paste0(data,"FigS9.pdf"), height = 3.4, width = 7, units = "in")
```

