---
title: "G x Sex eQTLs"
author: "Meng"
date: "2024-11-06"
output: html_document
---

```{r setup, include=FALSE}
library(arrow)
library(dplyr)
library(ggplot2)
library(patchwork)
library(tidyr)
data <- "/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/sex/"
```

# plot eQTLs showing significant G x Sex

```{r}
normalized_counts_ln <- read.table("/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/normalized_counts_ln_auto.bed", header = T, check.names = F)

# pval_adj_bh <- sex_inter_top %>% filter(pval_adj_bh < 0.1) # 5
pval_adj_bh <- read.csv(paste0(data, "sex_interaction_padj_bh01.csv"))
write.table(pval_adj_bh$variant_id, file = paste0(data, "pval_adj_bh.txt"), row.names = F, quote = F, col.names = F)

pval_adj_bh <- pval_adj_bh[order(pval_adj_bh$pval_adj_bh),]
pval_adj_bh[,c(1:2,18)]
```
  phenotype_id  variant_id pval_adj_bh
1 TX_paternal_00002596 1:159008371  0.00197039
2 TX_paternal_00010325  2:14033315  0.00197039
3 TX_paternal_00025295  3:10654019  0.00317137
4 TX_paternal_00001727  1:70371578  0.06471960
5 TX_paternal_00015022 2:250749626  0.07920580

## 1 TX_paternal_00002596 1:159008371 G	A
```{r}
expr <- normalized_counts_ln %>% filter(gene_id == "TX_paternal_00002596")
geno <- read.table(paste0(data,"A1_159008371.vcf"), header = T, check.names = F)
pheno <- expr[,c(5:147)]
geno1 <- geno[,c(5:147)]
eqtl <- rbind(geno1,pheno)
eqtl <- as.data.frame(t(eqtl))
eqtl <- eqtl %>% mutate(id = row.names(eqtl))
colnames(eqtl) <- c("genotype", "phenotype", "id")
eqtl$genotype <- factor(eqtl$genotype)
levels(eqtl$genotype)
table(eqtl$genotype)

levels(eqtl$genotype) <-c ("./.", "GG","GA", "AA")
eqtl <- eqtl %>% filter(genotype != "./." & genotype != "AA")
eqtl$phenotype <- as.numeric(eqtl$phenotype)

sex <- rep(NA, length(eqtl$id))
sex[grep("M", eqtl$id)] <- "M"
sex[grep("F", eqtl$id)] <- "F"
eqtl <- data.frame(eqtl, sex)
write.table(eqtl, file = paste0(data, "A1_159008371.eqtl"), quote = F)
#eqtl <- read.table(paste0(data, "A1_159008371.eqtl"), header = T)
p1 <- ggplot(eqtl, aes(x=genotype, y=phenotype, colour=sex)) + 
  geom_boxplot(position=position_dodge2(0.85)) + 
  geom_jitter(shape=16, position=position_jitterdodge(0.2), alpha=0.3) + 
  theme_classic() + theme(text = element_text(size = 10))+ scale_colour_manual(values = c("purple", "orange"))+labs(x="Genotype", y="Expression", title = "a")
p1 
```

## 2 TX_paternal_00010325  2:14033315 A C
```{r}
expr <- normalized_counts_ln %>% filter(gene_id == "TX_paternal_00010325")
geno <- read.table(paste0(data,"A2_14033315.vcf"), header = T, check.names = F)
pheno <- expr[,c(5:147)]
geno1 <- geno[,c(5:147)]
eqtl <- rbind(geno1,pheno)
eqtl <- as.data.frame(t(eqtl))
eqtl <- eqtl %>% mutate(id = row.names(eqtl))
colnames(eqtl) <- c("genotype", "phenotype", "id")
eqtl$genotype <- factor(eqtl$genotype)
levels(eqtl$genotype)
table(eqtl$genotype)

levels(eqtl$genotype) <-c ("./.", "AC", "CC")
eqtl <- eqtl %>% filter(genotype != "./.")
eqtl$phenotype <- as.numeric(eqtl$phenotype)

sex <- rep(NA, length(eqtl$id))
sex[grep("M", eqtl$id)] <- "M"
sex[grep("F", eqtl$id)] <- "F"
eqtl <- data.frame(eqtl, sex)
write.table(eqtl, file = paste0(data, "A2_14033315.eqtl"), quote = F)
#eqtl <- read.table(paste0(data, "A2_14033315.eqtl"), header = T)

p2 <- ggplot(eqtl, aes(x=genotype, y=phenotype, colour=sex)) + 
  geom_boxplot(position=position_dodge2(0.85)) + 
  geom_jitter(shape=16, position=position_jitterdodge(0.2), alpha=0.3) + 
  theme_classic() + theme(text = element_text(size = 10))+ scale_colour_manual(values = c("purple", "orange"))+labs(x="Genotype", y="Expression", title = "b")
p2 
```

## 3 TX_paternal_00025295  3:10654019 G A
```{r}
expr <- normalized_counts_ln %>% filter(gene_id == "TX_paternal_00025295")
geno <- read.table(paste0(data,"A3_10654019.vcf"), header = T, check.names = F)
pheno <- expr[,c(5:147)]
geno1 <- geno[,c(5:147)]
eqtl <- rbind(geno1,pheno)
eqtl <- as.data.frame(t(eqtl))
eqtl <- eqtl %>% mutate(id = row.names(eqtl))
colnames(eqtl) <- c("genotype", "phenotype", "id")
eqtl$genotype <- factor(eqtl$genotype)
levels(eqtl$genotype)
table(eqtl$genotype)

levels(eqtl$genotype) <-c ("./.", "GG","GA", "AA")
eqtl <- eqtl %>% filter(genotype != "./." & genotype != "GG")
eqtl$phenotype <- as.numeric(eqtl$phenotype)

sex <- rep(NA, length(eqtl$id))
sex[grep("M", eqtl$id)] <- "M"
sex[grep("F", eqtl$id)] <- "F"
eqtl <- data.frame(eqtl, sex)
write.table(eqtl, file = paste0(data, "A3_10654019.eqtl"), quote = F)
#eqtl <- read.table(paste0(data, "A3_10654019.eqtl"), header = T)

p3 <- ggplot(eqtl, aes(x=genotype, y=phenotype, colour=sex)) + 
  geom_boxplot(position=position_dodge2(0.85)) + 
  geom_jitter(shape=16, position=position_jitterdodge(0.2), alpha=0.3) + 
  theme_classic() + theme(text = element_text(size = 10))+ scale_colour_manual(values = c("purple", "orange"))+labs(x="Genotype", y="Expression", title = "c")
p3 
```

## 4 TX_paternal_00001727  1:70371578 G T
```{r}
expr <- normalized_counts_ln %>% filter(gene_id == "TX_paternal_00001727")
geno <- read.table(paste0(data,"A1_70371578.vcf"), header = T, check.names = F)
pheno <- expr[,c(5:147)]
geno1 <- geno[,c(5:147)]
eqtl <- rbind(geno1,pheno)
eqtl <- as.data.frame(t(eqtl))
eqtl <- eqtl %>% mutate(id = row.names(eqtl))
colnames(eqtl) <- c("genotype", "phenotype", "id")

eqtl$genotype <- factor(eqtl$genotype)
levels(eqtl$genotype)
levels(eqtl$genotype) <-c ("./.","GT", "TT")
eqtl <- eqtl %>% filter(genotype != "./.")
eqtl$phenotype <- as.numeric(eqtl$phenotype)

sex <- rep(NA, length(eqtl$id))
sex[grep("M", eqtl$id)] <- "M"
sex[grep("F", eqtl$id)] <- "F"
eqtl <- data.frame(eqtl, sex)
write.table(eqtl, file = paste0(data, "A1_70371578.eqtl"), quote = F)
#eqtl <- read.table(paste0(data, "A1_70371578.eqtl"), header = T)
p4 <- ggplot(eqtl, aes(x=genotype, y=phenotype, colour=sex)) + 
  geom_boxplot(position=position_dodge2(0.85)) + 
  geom_jitter(shape=16, position=position_jitterdodge(0.2), alpha=0.3) + 
  theme_classic() + theme(text = element_text(size = 10))+ scale_colour_manual(values = c("purple", "orange"))+labs(x="Genotype", y="Expression", title = "d")
p4 
```



## 5 TX_paternal_00015022 2:250749626 A G
```{r}
expr <- normalized_counts_ln %>% filter(gene_id == "TX_paternal_00015022")
geno <- read.table(paste0(data,"A2_250749626.vcf"), header = T, check.names = F)
pheno <- expr[,c(5:147)]
geno1 <- geno[,c(5:147)]
eqtl <- rbind(geno1,pheno)
eqtl <- as.data.frame(t(eqtl))
eqtl <- eqtl %>% mutate(id = row.names(eqtl))
colnames(eqtl) <- c("genotype", "phenotype", "id")
eqtl$genotype <- factor(eqtl$genotype)
levels(eqtl$genotype)
table(eqtl$genotype)

levels(eqtl$genotype) <-c ("./.","AG", "GG")
eqtl <- eqtl %>% filter(genotype != "./.")
eqtl$phenotype <- as.numeric(eqtl$phenotype)

sex <- rep(NA, length(eqtl$id))
sex[grep("M", eqtl$id)] <- "M"
sex[grep("F", eqtl$id)] <- "F"
eqtl <- data.frame(eqtl, sex)
write.table(eqtl, file = paste0(data, "A2_250749626.eqtl"), quote = F)
#eqtl <- read.table(paste0(data, "A2_250749626.eqtl"), header = T)

p5 <- ggplot(eqtl, aes(x=genotype, y=phenotype, colour=sex)) + 
  geom_boxplot(position=position_dodge2(0.85)) + 
  geom_jitter(shape=16, position=position_jitterdodge(0.2), alpha=0.3) + 
  theme_classic() + theme(text = element_text(size = 10))+ scale_colour_manual(values = c("purple", "orange"))+labs(x="Genotype", y="Expression", title = "e")
p5
```



## plot all
```{r}
p1 + p2 + p3 + p4 + p5 + plot_layout(nrow = 2,guides = "collect") & theme(legend.position = "right")
```


# check eQTLs that only show sex effect
4:79507645
```{r}
p_i_adj_01 %>% filter(variant_id == "4:79507645")
```
 phenotype_id variant_id start_distance       af ma_samples ma_count
1 TX_paternal_00030453 4:79507645           1521 0.135659         35       35
     pval_g      b_g   b_g_se      pval_i     b_i   b_i_se  pval_gi      b_gi
1 0.0901987 0.786358 0.460851 6.07478e-17 1.43296 0.149801 0.185892 -0.398788
   b_gi_se tests_emt pval_emt pval_adj_bh pval_gi_adj   pval_i_adj pval_g_adj
1 0.299964        23        1           1   0.2131096 5.907521e-14  0.1374715

## A4_79507645.vcf
ref C alt T
```{r}
normalized_counts_ln <- read.table("/Users/yuanmeng/Library/CloudStorage/OneDrive-UniversityofToronto/Manuscripts/rumex_eqtl/normalized_counts_ln_auto.bed", header = T, check.names = F)
expr <- normalized_counts_ln %>% filter(gene_id == "TX_paternal_00030453")
geno <- read.table(paste0(data,"A4_79507645.vcf"), header = T, check.names = F)

pheno <- expr[,c(5:147)]
geno1 <- geno[,c(5:147)]
eqtl <- rbind(geno1,pheno)
eqtl <- as.data.frame(t(eqtl))

eqtl <- eqtl %>% mutate(id = row.names(eqtl))
colnames(eqtl) <- c("genotype", "phenotype", "id")

eqtl$genotype <- factor(eqtl$genotype)
eqtl <- eqtl %>% filter(genotype != "./.")
levels(eqtl$genotype)
eqtl$phenotype <- as.numeric(eqtl$phenotype)

sex <- rep(NA, length(eqtl$id))
sex[grep("M", eqtl$id)] <- "M"
sex[grep("F", eqtl$id)] <- "F"
eqtl <- data.frame(eqtl, sex)

write.table(eqtl, file = paste0(data, "A4_79507645.eqtl"), quote = F)

f <- eqtl %>% filter(sex == "F")
m <- eqtl %>% filter(sex == "M")
table(f$genotype)
table(m$genotype)
```
./.  CC C/T 
  0  49  15
./.  CC C/T 
  0  45  20 
```{r}
eqtl <- eqtl %>% filter(genotype == "0/0" | genotype == "0/1")
levels(eqtl$genotype) <-c ("./.","CC", "CT")
# p1 <- ggplot(eqtl, aes(x=sex, y=phenotype, colour=genotype)) + 
#   geom_boxplot(position=position_dodge2(0.85)) + 
#   geom_jitter(shape=16, position=position_jitterdodge(0.2), alpha=0.3) + 
#   theme_classic()+ theme(text = element_text(size = 15))

p2 <- ggplot(eqtl, aes(x=genotype, y=phenotype, colour=sex)) + 
  geom_boxplot(position=position_dodge2(0.85)) + 
  geom_jitter(shape=16, position=position_jitterdodge(0.2), alpha=0.3) + 
  theme_classic() + theme(text = element_text(size = 18))+ scale_colour_manual(values = c("red", "blue"))+labs(x="Genotype", y="Expression")
p2 
```





